---
title: "White Oak Population Forecast"
resource_files:
- dashboard/states-shape/cb_2016_us_state_500k.shp
- dashboard/unit_bnd/eastunits.shp
- cb_2016_us_state_500k.shp
- eastunits.shp
- cb_2016_us_state_500k.cpg
- cb_2016_us_state_500k.dbf
- cb_2016_us_state_500k.prj
- cb_2016_us_state_500k.shp
- cb_2016_us_state_500k.shp.ea.iso.xml
- cb_2016_us_state_500k.shp.iso.xml
- cb_2016_us_state_500k.shp.xml
- cb_2016_us_state_500k.shx
- eastunits.dbf
- eastunits.shp
- eastunits.shp.xml
- eastunits.shx
- forestry_divisions.cpg
- forestry_divisions.dbf
- forestry_divisions.prj
- forestry_divisions.shp
- forestry_divisions.shx
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    logo: bf-logo.png
    orientation: rows
    source_code: embed
    theme: readable
    vertical_layout: fill
---

```{r}
# Change the color of the title bar and the input sidebar. Remove the background from the leaflet map.

```

 <style>                     
.navbar {
  background-color: #C0C0C0;
}

.section.sidebar {
  background-color: #DCDCDC;
}

.leaflet-container {
    background-color:rgba(255,0,0,0.0);
}
</style> 

```{r setup, include=FALSE}
# Load required libraries
library(flexdashboard)
library(tidyverse)
library(ggplot2)
library(shiny)
library(dygraphs)
library(plotly)
library(formattable)
library(leaflet)
library(rgdal)

```
TESTING
======================================================================
```{r}
# Load required libraries
library(flexdashboard)
library(tidyverse)
library(ggplot2)
library(shiny)
library(plotly)
library(formattable)
library(leaflet)
library(rgdal)

# Read in the data. This data was cleaned and merged in the for_dashboard.R script.
data <- read.csv("WhiteOakData/white-oak-clean.csv")

data[is.na(data)] <- 0

# Aggregate the growth, removal, and mortality data to the country level
national <- data %>% 
  group_by(EVAL_GRP_YEAR) %>%
  summarise(total.growth = sum(SELECT_WHITE_OAK_GROWTH, na.rm = TRUE),
            total.mortality = sum(SELECT_WHITE_OAK_MORTALITY, na.rm = TRUE),
            total.removal = sum(SELECT_WHITE_OAK_REMOVAL, na.rm = TRUE))

# Read in and manipulate inventory data. These are the raw files from the US Forest service.
inv23 <- read_csv("WhiteOakData/white_oak_inv_23.csv")
inv24 <- read_csv("WhiteOakData/white_oak_inv_24.csv")
inv33 <- read_csv("WhiteOakData/white_oak_inv_33.csv")

inventory <- rbind(inv23, inv24, inv33)

inventory[is.na(inventory)] <- 0

# Create an identifier to merge with the shape files later.
inventory$UNIT <- paste0(inventory$STATECD, "0", inventory$UNITCD)

# Don't think I need both of these next two
inventory.state <- inventory[,c(3,5,6,10:27)] %>%
  group_by(STATE_ABBR, EVAL_GRP_YEAR, UNITCD) %>% 
  summarise_all(funs(sum(., na.rm = TRUE)))

inv.state.rolled <- inventory.state[,c(1,2,4:21)] %>% 
  group_by(STATE_ABBR, EVAL_GRP_YEAR) %>%
  summarise_all(funs(sum(., na.rm = TRUE)))

# Aggregate to unit level inventory for heat map.
inv.unit.rolled <- inventory[,c(3,5,31,10:27)] %>% 
  group_by(STATE_ABBR, EVAL_GRP_YEAR, UNIT) %>% 
  summarise_all(funs(sum(., na.rm = TRUE)))

# Create list of unique states in the growth/removal/mortality data and filter the inventory data so that all states appear in both sets.
state_list <- unique(data$STATE_ABBR)

inv_national <- inv.state.rolled %>%
  filter(STATE_ABBR %in% state_list) 

# Aggregate the inventory data to the country level.
inv_national <- inv_national[2:20] %>% 
  group_by(EVAL_GRP_YEAR) %>% 
  summarise_all(funs(sum(., na.rm = TRUE)))

# Select relevant columns and merge the national inventory data with the national growth/removal/mortality data.
inventory_national <- inv_national %>% 
  select(EVAL_GRP_YEAR, SELECT_WHITE_OAK_INV_CUFT_GS, SELECT_WHITE_OAK_INV_CUFT_LIVE)

national <- merge(national, inventory_national, by = "EVAL_GRP_YEAR")

colnames(national)[5:6] <- c("inv.gs", "inv.live")

national$growth.pct <- 0
national$mortality.pct <- 0
national$removal.pct <- 0

# Calculate the growth/removal/mortality percent of growing stock inventory.
national$growth.pct <- national$total.growth/national$inv.gs
national$mortality.pct <- national$total.mortality/national$inv.gs
national$removal.pct <- national$total.removal/national$inv.gs
  
# Exclude data prior to 1991 and after 2014. All units were not sampled prior to 1991, and '15, `16 and `17 data may be incomplete per US Forest Service.
national <- filter(national, EVAL_GRP_YEAR <= 2014 & EVAL_GRP_YEAR >= 1991)

# Build models. I regressed against the values 1:24 so I could verify the intercept. Whether you regress against this or the actual year values, the slope remains the same.
model_growth <- lm(growth.pct ~ c(1:24), data = national)
model_mortality <- lm(mortality.pct ~ c(1:24), data = national)
model_removal <- lm(removal.pct ~ c(1:24), data = national)

# Store slope coefficients
growth_slope <- coef(model_growth)[[2]]
mortality_slope <- coef(model_mortality)[[2]]
removal_slope <- coef(model_removal)[[2]]

# Store upper and lower bounds on slope coefficients
growth_slope_upper <- confint(model_growth)[4]
growth_slope_lower <- confint(model_growth)[2]

mortality_slope_upper <- confint(model_mortality)[4]
mortality_slope_lower <- confint(model_mortality)[2]

removal_slope_upper <- confint(model_removal)[4]
removal_slope_lower <- confint(model_removal)[2]

# Use average of 2010 - 2014 as intercept
growth_int <- mean(national$growth.pct[20:24])
mortality_int <- mean(national$mortality.pct[20:24])
removal_int <- mean(national$removal.pct[20:24])
```

Forecast
=======================================================================

Inputs {.sidebar}
-----------------------------------------------------------------------

\

\


\


\

\

*See **Documentation** for explanation of forecast methods.*

Row
-----------------------------------------------------------------------

### White Oak Inventory Forecast *with 95% Confidence Interval*

```{r echo = F,warning=FALSE,message=FALSE,error=FALSE}
# Build "forecast" dataframe
future_alt <- NULL
future_alt$year <- 2014:2064
future_alt <- data.frame(future_alt)

future_alt$inventory <- 0
future_alt$inventory.high <- 0
future_alt$inventory.low <- 0
future_alt$growth <- 0
future_alt$growth.high <- 0
future_alt$growth.low <- 0
future_alt$removals <- 0
future_alt$removal.low <- 0
future_alt$removal.high <- 0
future_alt$mortality <- 0
future_alt$mortality.high <- 0
future_alt$mortality.low <- 0
future_alt$urbanization <- 0
future_alt$urbanization.high <- 0
future_alt$urbanization.low <- 0
future_alt$demand <- 0
future_alt$`demand % of inventory` <- 0
future_alt$`demand % of removals` <- 0

# Fill in data for 2014
inv_14 <- inventory_national %>% 
  filter(EVAL_GRP_YEAR == 2014) %>% 
  select(SELECT_WHITE_OAK_INV_CUFT_GS)

future_alt$inventory[1] <- as.numeric(inv_14[1,1])
future_alt$inventory.high[1] <- as.numeric(inv_14[1,1])
future_alt$inventory.low[1] <- as.numeric(inv_14[1,1])
future_alt$growth[1] <- national$total.growth[nrow(national)]
future_alt$removal.high[1] <- national$total.removal[nrow(national)]
future_alt$removal.low[1] <- national$total.removal[nrow(national)]
future_alt$removals[1] <- national$total.removal[nrow(national)]
future_alt$mortality[1] <- national$total.mortality[nrow(national)]
future_alt$growth.high[1] <- national$total.growth[nrow(national)]
future_alt$mortality.high[1] <- national$total.mortality[nrow(national)]
future_alt$growth.low[1] <- national$total.growth[nrow(national)]
future_alt$mortality.low[1] <- national$total.mortality[nrow(national)]

# Calculate growth and mortality % "forecast"
future_alt$dummy <- 0:50

future_alt$new_growth <- ((future_alt$dummy*growth_slope)+growth_int)
future_alt$new_mortality <- ((future_alt$dummy*mortality_slope)+mortality_int)
future_alt$new_removal <- ((future_alt$dummy*removal_slope)+removal_int)

future_alt$high_growth <- ((future_alt$dummy*growth_slope_upper)+growth_int)
future_alt$high_mortality <- ((future_alt$dummy*mortality_slope_upper)+mortality_int)
future_alt$high_removal <- ((future_alt$dummy*removal_slope_upper)+removal_int)

future_alt$low_growth <- ((future_alt$dummy*growth_slope_lower)+growth_int)
future_alt$low_mortality <- ((future_alt$dummy*mortality_slope_lower)+mortality_int)
future_alt$low_removal <- ((future_alt$dummy*removal_slope_lower)+removal_int)

# growth, removal, mortality cannot be a negative % of inventory
future_alt$low_removal <- ifelse(future_alt$low_removal <= 0, 0, future_alt$low_removal)
future_alt$low_growth <- ifelse(future_alt$low_growth <= 0, 0, future_alt$low_growth)
future_alt$low_mortality <- ifelse(future_alt$low_mortality <= 0, 0, future_alt$low_mortality)

future_alt$high_removal <- ifelse(future_alt$high_removal <= 0, 0, future_alt$high_removal)
future_alt$high_growth <- ifelse(future_alt$high_growth <= 0, 0, future_alt$high_growth)
future_alt$high_mortality <- ifelse(future_alt$high_mortality <= 0, 0, future_alt$high_mortality)

future_alt$new_removal <- ifelse(future_alt$new_removal <= 0, 0, future_alt$new_removal)
future_alt$new_growth <- ifelse(future_alt$new_growth <= 0, 0, future_alt$new_growth)
future_alt$new_mortality <- ifelse(future_alt$new_mortality <= 0, 0, future_alt$new_mortality)

future_alt$dummy <- NULL

# renderPlot for inputs to work
# renderggiraph for ggiraph graph
# renderPlotly for plotly graph

renderPlotly({
  
  # fill in forecast table row by row
  for (i in 2:nrow(future_alt)) {
    # Inventory = Inventory + Growth - Removal - Mortality - Urbanization  
    future_alt$inventory[i] <- future_alt$inventory[i-1] + future_alt$growth[i-1] - future_alt$removals[i-1] - future_alt$mortality[i-1] - future_alt$urbanization[i-1]
    
    # Upper Bound Inventory = Upper Bound Inventory + High Growth - Low Removal - Low Mortality - High Urbanization (i.e. the Urbanization calculated from the High Inventory)
    future_alt$inventory.high[i] <- future_alt$inventory.high[i-1] + future_alt$growth.high[i-1] - future_alt$removal.low[i-1] - future_alt$mortality.low[i-1] - future_alt$urbanization.high[i-1]
    
    # Lower Bound Inventory = Lower Bound Inventory + Low Growth - High Removal - High Mortality - Low Urbanization (i.e. the Urbanization calculated from the Low Inventory)
    future_alt$inventory.low[i] <- future_alt$inventory.low[i-1] + future_alt$growth.low[i-1] - future_alt$removal.high[i-1] - future_alt$mortality.high[i-1] - future_alt$urbanization.low[i-1]
    
    future_alt$growth[i] <- future_alt$inventory[i]*(future_alt$new_growth[i])    
    future_alt$growth.high[i] <- future_alt$inventory.high[i]*(future_alt$high_growth[i])
    future_alt$growth.low[i] <- future_alt$inventory.low[i]*(future_alt$low_growth[i])
    
    future_alt$mortality[i] <- future_alt$inventory[i]*(future_alt$new_mortality[i])
    future_alt$mortality.high[i] <- future_alt$inventory.low[i]*(future_alt$high_mortality[i])
    future_alt$mortality.low[i] <- future_alt$inventory.high[i]*(future_alt$low_mortality[i])
    
    future_alt$removals[i] <- future_alt$inventory[i]*future_alt$new_removal[i]
    future_alt$removal.high[i] <- future_alt$inventory.low[i]*future_alt$high_removal[i]
    future_alt$removal.low[i] <- future_alt$inventory.high[i]*future_alt$low_removal[i]
    
    future_alt$urbanization[i] <- future_alt$inventory[i]*(as.numeric(input$urbanization2)/100)
    future_alt$urbanization.high[i] <- future_alt$inventory.high[i]*(as.numeric(input$urbanization2)/100)
    future_alt$urbanization.low[i] <- future_alt$inventory.low[i]*(as.numeric(input$urbanization2)/100)
    
  }
  
  future_alt$inventory <- round(as.numeric(future_alt$inventory),0)
  future_alt$inventory.high <- round(as.numeric(future_alt$inventory.high),0)
  future_alt$inventory.low <- round(as.numeric(future_alt$inventory.low),0)  
  
  # Select relevant columns for inventory plot.
  future_plt <- select(future_alt, year, inventory, inventory.high, inventory.low)
  
  # Transform to long format for easier ggplot use.
  future_plt <- tidyr::gather(future_plt, 
                              key = 'type', 
                              value = 'forecast', 
                              -year)
  
  # Divide by 1 Billion for cleaner y-axis
  future_plt$forecast_div <- future_plt$forecast/1000000000
  
  # Bounds as factors for color differential
  future_plt$bounds <- ifelse(future_plt$type == "inventory", "n", "y")
  future_plt$bounds <- factor(future_plt$bounds, levels = c("y", "n"))
  
  # Build plot
  plt_alt <- ggplot(data = future_plt, aes(x = year, y = forecast_div, linetype = bounds, colour = bounds, size = bounds, group = type)) + 
    geom_line(aes(text = paste0("Year: ", year, "<br>Forecast: ", formattable::comma(forecast)))) +
    scale_colour_manual(values = c("maroon", "darkgreen")) +
    scale_linetype_manual(values = c(3,1)) +
    scale_size_manual(values = c(1,1)) +
    theme(panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          legend.position = 'none') +
    ylab("Inventory (Billions, Cubic Feet)") +
    xlab("") +
    scale_y_continuous(limits = c(0,110))
  
  
  ggplotly(plt_alt, tooltip = 'text')
  
})
```


Row
-----------------------------------------------------------------------

### Current Inventory (2014) by FIA Unit

```{r echo = F,warning=FALSE,message=FALSE,error=FALSE}
# Aggregate to state level to merge with state shape file
inventory.map <- inv.state.rolled %>%
  select(STATE_ABBR, EVAL_GRP_YEAR, SELECT_WHITE_OAK_INV_CUFT_GS) %>%
  group_by(STATE_ABBR, EVAL_GRP_YEAR) %>%
  summarise(tot = sum(SELECT_WHITE_OAK_INV_CUFT_GS, na.rm = TRUE)) %>%
  filter(EVAL_GRP_YEAR == 2014)

# Aggregate to unit level to merge with unit shape file
inventory.map.unit <- inv.unit.rolled %>%
  select(STATE_ABBR, EVAL_GRP_YEAR, UNIT, SELECT_WHITE_OAK_INV_CUFT_GS) %>%
  group_by(STATE_ABBR, EVAL_GRP_YEAR, UNIT) %>%
  summarise(tot = sum(SELECT_WHITE_OAK_INV_CUFT_GS)) %>%
  filter(EVAL_GRP_YEAR == 2014)

colnames(inventory.map)[1] <- "STUSPS"

# Read in state shape file
states_shp <- readOGR("MapFiles", layer = "cb_2016_us_state_500k",
                      GDAL1_integer64_policy = TRUE, verbose = FALSE)

# Remove non-contiguous states/territories
states.shp <- subset(states_shp, !states_shp$STUSPS %in% c("AK", "GU", "MP", "VI", "AS", "PR", "HI"))

# Merge shape and inventory files
states.shp <- sp::merge(states.shp, inventory.map, by = "STUSPS")

stusps <- as.vector(as.character(states.shp@data$STUSPS))

# Read in unit shape file (from US forest service)
units_shp <- readOGR("MapFiles", layer = "eastunits",
                     GDAL1_integer64_policy = TRUE, verbose = FALSE)

# Merge inventory and shape files. Log transform for a more distinct break in heat map colors. 
units.shp <- sp::merge(units_shp, inventory.map.unit, by = "UNIT")
units.shp@data <- units.shp@data[complete.cases(units.shp@data),]
units.shp@data$tot[units.shp@data$tot == 0] <- min(units.shp@data$tot[units.shp@data$tot > 0])-5000
units.shp@data$tot.log <- log(units.shp@data$tot)

# Build map
output$inventory_map_rm <- renderLeaflet({
  
  # Set id to state abbreviation
  id <- as.vector(states.shp@data$STUSPS)
  
  leaflet(units.shp, options = leafletOptions(zoomControl = FALSE))%>%
    addPolygons(color = '#444444', weight = 1, smoothFactor = 0.5,
                opacity = 1.0, fillOpacity = 0.5,
                fillColor = ~colorNumeric("YlGn", tot.log)(tot.log)) %>%
    # Layer transparent state shapes over units for clickable states
    addPolygons(data = states.shp, layerId = id, color = '#000000', weight = 1, smoothFactor = 0.5,
                opacity = 1.0, fillOpacity = 0, label = id,
                highlightOptions = highlightOptions(color = "white", weight = 2,
                                                    bringToFront = TRUE))
  
})

leafletOutput('inventory_map_rm')
```


### Inventory Forecast by State *with 95% Confidence Interval*

```{r echo = F,warning=FALSE,message=FALSE,error=FALSE}
data$STATE_ABBR <- as.character(data$STATE_ABBR)

# renderPlotly for inputs to work
renderPlotly({
  # Store the id from the map click
  st.alt <- as.character(input$inventory_map_rm_shape_click$id)
  
  # Validate statement to return a helpful error if they haven't yet clicked on a state.
  # Filter data to the selected state.  
  validate(need(try(state.data.alt <-
                      filter(data, STATE_ABBR == st.alt))
                , "Click on a state to the left to view the state-level forecast."))
  
  state.data.alt <- state.data.alt[, c(1, 2, 5:7)] %>%
    group_by(STATE_ABBR, EVAL_GRP_YEAR) %>%
    summarise_all(funs(sum(., na.rm = TRUE)))
  
  state.inv.alt <- inv.state.rolled %>%
    select(STATE_ABBR, EVAL_GRP_YEAR, SELECT_WHITE_OAK_INV_CUFT_GS) %>%
    group_by(STATE_ABBR, EVAL_GRP_YEAR) %>%
    summarise(tot = sum(SELECT_WHITE_OAK_INV_CUFT_GS, na.rm = TRUE))
  
  state.mod.alt <-
    merge(state.data.alt, state.inv.alt, by = c("STATE_ABBR", "EVAL_GRP_YEAR"))
  
  colnames(state.mod.alt)[3:6] <-
    c("total.growth",
      "total.mortality",
      "total.removal",
      "inv.gs")
  
  state.mod.alt$growth.change <- state.mod.alt$total.growth / state.mod.alt$inv.gs
  state.mod.alt$mortality.change <- state.mod.alt$total.mortality / state.mod.alt$inv.gs
  state.mod.alt$removal.change <- state.mod.alt$total.removal / state.mod.alt$inv.gs
  
  state.mod.alt <-
    filter(state.mod.alt, EVAL_GRP_YEAR <= 2014 & EVAL_GRP_YEAR >= 1991)
  
  #check for consecutive years
  validate(need(try(state.mod.alt$conesc <- 2), "There is no data associated with the selected state. Please select a colored state on the map to view the state level forecast."))
  
  for(i in 1:(nrow(state.mod.alt)-1)) {
    if((state.mod.alt$EVAL_GRP_YEAR[i+1] == (state.mod.alt$EVAL_GRP_YEAR[i]) + 1)) {
      state.mod.alt$consec[i] <- 0
    } else {
      state.mod.alt$consec[i] <- 1
    }
  }
  
  state.mod.alt$consec[nrow(state.mod.alt)] <- 0 
  
  state.mod.alt <- filter(state.mod.alt, consec == 0)
  
  # Build model
  num.alt <- nrow(state.mod.alt)
  
  # Validate statement to return a helpful error if the selected state does not have associated data.
  validate(need(try(model_growth_st <- lm(growth.change ~ c(1:num.alt), data = state.mod.alt))
                , "There is no forecast available for selected state."))
  model_mortality_st <- lm(mortality.change ~ c(1:num.alt), data = state.mod.alt)
  model_removal_st <- lm(removal.change ~ c(1:num.alt), data = state.mod.alt)
  
  # Store coefficients
  growth_slope_st <- coef(model_growth_st)[[2]]
  mortality_slope_st <- coef(model_mortality_st)[[2]]
  removal_slope_st <- coef(model_removal_st)[[2]]
  
  # Store upper and lower bounds
  growth_slope_upper_st <- confint(model_growth_st)[4]
  growth_slope_lower_st <- confint(model_growth_st)[2]
  
  mortality_slope_upper_st <- confint(model_mortality_st)[4]
  mortality_slope_lower_st <- confint(model_mortality_st)[2]
  
  removal_slope_upper_st <- confint(model_removal_st)[4]
  removal_slope_lower_st <- confint(model_removal_st)[2]
  
  # Use average of 2010 - 2014 as intercept
  for_int <- filter(state.mod.alt, EVAL_GRP_YEAR >= 2010)
  growth_int_st <- mean(for_int$growth.change)
  mortality_int_st <- mean(for_int$mortality.change)
  removal_int_st <- mean(for_int$removal.change)
  
  # Build "forecast" dataframe
  future_st_alt <- NULL
  future_st_alt$year <- 2014:2040
  future_st_alt <- data.frame(future_st_alt)
  
  future_st_alt$inventory <- 0
  future_st_alt$inventory.high <- 0
  future_st_alt$inventory.low <- 0
  future_st_alt$growth <- 0
  future_st_alt$growth.high <- 0
  future_st_alt$growth.low <- 0
  future_st_alt$removals <- 0
  future_st_alt$removal.low <- 0
  future_st_alt$removal.high <- 0
  future_st_alt$mortality <- 0
  future_st_alt$mortality.high <- 0
  future_st_alt$mortality.low <- 0
  future_st_alt$urbanization <- 0
  future_st_alt$urbanization.high <- 0
  future_st_alt$urbanization.low <- 0
  
  # Fill in data for 2014
  inv_14 <- state.mod.alt %>%
    filter(EVAL_GRP_YEAR == 2014) %>%
    select(inv.gs)
  
  future_st_alt$inventory[1] <- as.numeric(inv_14[1, 1])
  future_st_alt$inventory.high[1] <- as.numeric(inv_14[1, 1])
  future_st_alt$inventory.low[1] <- as.numeric(inv_14[1, 1])
  future_st_alt$growth[1] <- state.mod.alt$total.growth[nrow(state.mod.alt)]
  future_st_alt$removals[1] <- state.mod.alt$total.removal[nrow(state.mod.alt)]
  future_st_alt$removal.low[1] <- state.mod.alt$total.removal[nrow(state.mod.alt)]
  future_st_alt$removal.high[1] <- state.mod.alt$total.removal[nrow(state.mod.alt)]
  future_st_alt$mortality[1] <- state.mod.alt$total.mortality[nrow(state.mod.alt)]
  future_st_alt$growth.high[1] <- state.mod.alt$total.growth[nrow(state.mod.alt)]
  future_st_alt$mortality.high[1] <- state.mod.alt$total.mortality[nrow(state.mod.alt)]
  future_st_alt$growth.low[1] <- state.mod.alt$total.growth[nrow(state.mod.alt)]
  future_st_alt$mortality.low[1] <- state.mod.alt$total.mortality[nrow(state.mod.alt)]
  
  # Calculate growth and mortality % "forecast"
  future_st_alt$dummy <- 0:26
  
  future_st_alt$new_growth <- ((future_st_alt$dummy * growth_slope_st) + growth_int_st)
  future_st_alt$new_mortality <- ((future_st_alt$dummy * mortality_slope_st) + mortality_int_st)
  future_st_alt$new_removal <- (future_st_alt$dummy * removal_slope_st) + removal_int_st
  
  future_st_alt$high_growth <- ((future_st_alt$dummy * growth_slope_upper_st) + growth_int_st)
  future_st_alt$high_mortality <- ((future_st_alt$dummy * mortality_slope_upper_st) + mortality_int_st)
  future_st_alt$high_removal <- (future_st_alt$dummy * removal_slope_upper_st) + removal_int_st
  
  future_st_alt$low_growth <- ((future_st_alt$dummy * growth_slope_lower_st) + growth_int_st)
  future_st_alt$low_mortality <- ((future_st_alt$dummy * mortality_slope_lower_st) + mortality_int_st)
  future_st_alt$low_removal <- (future_st_alt$dummy * removal_slope_lower_st) + removal_int_st
  
  # growth, removal, mortality cannot be a negative % of inventory
  future_st_alt$low_removal <- ifelse(future_st_alt$low_removal <= 0, 0, future_st_alt$low_removal)
  future_st_alt$low_growth <- ifelse(future_st_alt$low_growth <= 0, 0, future_st_alt$low_growth)
  future_st_alt$low_mortality <- ifelse(future_st_alt$low_mortality <= 0, 0, future_st_alt$low_mortality)
  
  future_st_alt$high_removal <- ifelse(future_st_alt$high_removal <= 0, 0, future_st_alt$high_removal)
  future_st_alt$high_growth <- ifelse(future_st_alt$high_growth <= 0, 0, future_st_alt$high_growth)
  future_st_alt$high_mortality <- ifelse(future_st_alt$high_mortality <= 0, 0, future_st_alt$high_mortality)
  
  future_st_alt$new_removal <- ifelse(future_st_alt$new_removal <= 0, 0, future_st_alt$new_removal)
  future_st_alt$new_growth <- ifelse(future_st_alt$new_growth <= 0, 0, future_st_alt$new_growth)
  future_st_alt$new_mortality <- ifelse(future_st_alt$new_mortality <= 0, 0, future_st_alt$new_mortality)
  
  future_st_alt$dummy <- NULL
  
  # fill in forecast table row by row
  for (i in 2:nrow(future_st_alt)) {
    future_st_alt$inventory[i] <-
      future_st_alt$inventory[i - 1] + future_st_alt$growth[i - 1] - future_st_alt$removals[i - 1] - future_st_alt$mortality[i - 1] - future_st_alt$urbanization[i - 1]
    
    future_st_alt$inventory.high[i] <-
      future_st_alt$inventory.high[i - 1] + future_st_alt$growth.high[i - 1] - future_st_alt$removal.low[i - 1] - future_st_alt$mortality.low[i - 1] - future_st_alt$urbanization.high[i - 1]
    
    future_st_alt$inventory.low[i] <-
      future_st_alt$inventory.low[i - 1] + future_st_alt$growth.low[i - 1] - future_st_alt$removal.high[i - 1] - future_st_alt$mortality.high[i - 1] - future_st_alt$urbanization.low[i - 1]
    
    future_st_alt$growth[i] <- future_st_alt$inventory[i] * (future_st_alt$new_growth[i])
    future_st_alt$growth.high[i] <- future_st_alt$inventory.high[i] * (future_st_alt$high_growth[i])
    future_st_alt$growth.low[i] <- future_st_alt$inventory.low[i] * (future_st_alt$low_growth[i])      
    
    future_st_alt$mortality[i] <- future_st_alt$inventory[i] * (future_st_alt$new_mortality[i])
    future_st_alt$mortality.high[i] <- future_st_alt$inventory.low[i] * (future_st_alt$high_mortality[i])
    future_st_alt$mortality.low[i] <- future_st_alt$inventory.high[i] * (future_st_alt$low_mortality[i])
    
    future_st_alt$removals[i] <- future_st_alt$inventory[i] * (future_st_alt$new_removal[i])
    future_st_alt$removal.high[i] <- future_st_alt$inventory.low[i] * future_st_alt$high_removal[i]
    future_st_alt$removal.low[i] <- future_st_alt$inventory.high[i] * future_st_alt$low_removal[i]
    
    future_st_alt$urbanization[i] <- future_st_alt$inventory[i] * (as.numeric(input$urbanization2) / 100)
    future_st_alt$urbanization.high[i] <- future_st_alt$inventory.high[i] * (as.numeric(input$urbanization2) / 100)
    future_st_alt$urbanization.low[i] <- future_st_alt$inventory.low[i] * (as.numeric(input$urbanization2) / 100)
    
  }
  
  future_st_alt$inventory <- round(as.numeric(future_st_alt$inventory),0)
  future_st_alt$inventory.high <- round(as.numeric(future_st_alt$inventory.high),0)
  future_st_alt$inventory.low <- round(as.numeric(future_st_alt$inventory.low),0)    
  
  future_plt <- select(future_st_alt, year, inventory, inventory.high, inventory.low)
  
  future_plt <- tidyr::gather(future_plt, 
                              key = 'type', 
                              value = 'forecast', 
                              -year)
  
  future_plt$forecast_div <- future_plt$forecast/1000000000
  future_plt$bounds <- ifelse(future_plt$type == "inventory", "n", "y")
  
  future_plt$bounds <- factor(future_plt$bounds, levels = c("y", "n"))
  
  plt_st_alt <- ggplot(data = future_plt, aes(x = year, y = forecast_div, linetype = bounds, colour = bounds, size = bounds, group = type)) + 
    geom_line(aes(text = paste0("Year: ", year, "<br>Forecast: ", formattable::comma(forecast)))) +
    scale_colour_manual(values = c("maroon", "darkgreen")) +
    scale_linetype_manual(values = c(3,1)) +
    scale_size_manual(values = c(1,1)) +
    theme(panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          legend.position = 'none') +
    ylab(paste0(unique(state.mod.alt$STATE_ABBR)," Inventory (Billions, Cubic Feet)")) +
    xlab("")
  
  
  ggplotly(plt_st_alt, tooltip = 'text')
  
})
```

Alternate Forecast
=======================================================================

Inputs {.sidebar}
-----------------------------------------------------------------------

\

\

Use the sliding bars below to adjust the increase in Annual Removal and Urbanization to see the effect these factors can have on the White Oak Population.

\

```{r echo = F,warning=FALSE,message=FALSE,error=FALSE}
tags$style(HTML(".sidebarPanel .body {background-color: lightgrey;}"))

tags$style(HTML(".js-irs-0 .irs-single, .js-irs-0 .irs-bar-edge, .js-irs-0 .irs-bar {background: green}"))
tags$style(HTML(".js-irs-1 .irs-single, .js-irs-1 .irs-bar-edge, .js-irs-1 .irs-bar {background: green}"))

sliderInput("removal", label = "Annual Removal Increase (%):",
            min = 0.0, max = 1.0, value = 0.5, step = 0.1)

sliderInput("urbanization", label = "Annual Urbanization (% of White Oak Inventory):",
            min = 0.0, max = 1.0, value = 0.1, step = 0.1)
```

\

\

*See **Documentation** for explanation of forecast methods.*

Row
-----------------------------------------------------------------------

### White Oak Inventory Forecast *with 95% Confidence Interval*

```{r echo = F,warning=FALSE,message=FALSE,error=FALSE}
# Build "forecast" dataframe
future <- NULL
future$year <- 2014:2064
future <- data.frame(future)

future$inventory <- 0
future$inventory.high <- 0
future$inventory.low <- 0
future$growth <- 0
future$growth.high <- 0
future$growth.low <- 0
future$removals <- 0
future$mortality <- 0
future$mortality.high <- 0
future$mortality.low <- 0
future$urbanization <- 0
future$urbanization.high <- 0
future$urbanization.low <- 0
future$demand <- 0
future$`demand % of inventory` <- 0
future$`demand % of removals` <- 0

# Fill in data for 2014
inv_14 <- inventory_national %>% 
  filter(EVAL_GRP_YEAR == 2014) %>% 
  select(SELECT_WHITE_OAK_INV_CUFT_GS)

future$inventory[1] <- as.numeric(inv_14[1,1])
future$inventory.high[1] <- as.numeric(inv_14[1,1])
future$inventory.low[1] <- as.numeric(inv_14[1,1])
future$growth[1] <- national$total.growth[nrow(national)]
future$removals[1] <- national$total.removal[nrow(national)]
future$mortality[1] <- national$total.mortality[nrow(national)]
future$growth.high[1] <- national$total.growth[nrow(national)]
future$mortality.high[1] <- national$total.mortality[nrow(national)]
future$growth.low[1] <- national$total.growth[nrow(national)]
future$mortality.low[1] <- national$total.mortality[nrow(national)]

# Calculate growth and mortality % "forecast"
future$dummy <- 0:50

future$new_growth <- ((future$dummy*growth_slope)+growth_int)
future$new_mortality <- ((future$dummy*mortality_slope)+mortality_int)

future$high_growth <- ((future$dummy*growth_slope_upper)+growth_int)
future$high_mortality <- ((future$dummy*mortality_slope_upper)+mortality_int)

future$low_growth <- ((future$dummy*growth_slope_lower)+growth_int)
future$low_mortality <- ((future$dummy*mortality_slope_lower)+mortality_int)

# Cannot be negative % of inventory
future$low_growth <- ifelse(future$low_growth <= 0, 0, future$low_growth)
future$low_mortality <- ifelse(future$low_mortality <= 0, 0, future$low_mortality)

future$high_growth <- ifelse(future$high_growth <= 0, 0, future$high_growth)
future$high_mortality <- ifelse(future$high_mortality <= 0, 0, future$high_mortality)

future$new_growth <- ifelse(future$new_growth <= 0, 0, future$new_growth)
future$new_mortality <- ifelse(future$new_mortality <= 0, 0, future$new_mortality)

future$dummy <- NULL

# renderPlot for inputs to work
# renderggiraph for ggiraph graph
# renderPlotly for plotly graph

renderPlotly({
  
  # fill in forecast table row by row
  for (i in 2:nrow(future)) {
    future$inventory[i] <- future$inventory[i-1] + future$growth[i-1] - future$removals[i-1] - future$mortality[i-1] - future$urbanization[i-1]
    
    future$inventory.high[i] <- future$inventory.high[i-1] + future$growth.high[i-1] - future$removals[i-1] - future$mortality.low[i-1] - future$urbanization.high[i-1]
    
    future$inventory.low[i] <- future$inventory.low[i-1] + future$growth.low[i-1] - future$removals[i-1] - future$mortality.high[i-1] - future$urbanization.low[i-1]
    
    future$growth[i] <- future$inventory[i]*(future$new_growth[i])
    future$growth.high[i] <- future$inventory.high[i]*(future$high_growth[i])
    future$growth.low[i] <- future$inventory.low[i]*(future$low_growth[i])
    
    future$mortality[i] <- future$inventory[i]*(future$new_mortality[i])
    future$mortality.high[i] <- future$inventory.low[i]*(future$high_mortality[i])
    future$mortality.low[i] <- future$inventory.high[i]*(future$low_mortality[i])
    
    future$removals[i] <- future$removals[i-1]*((1 + as.numeric(input$removal)/100))
    
    future$urbanization[i] <- future$inventory[i]*(as.numeric(input$urbanization)/100)
    future$urbanization.high[i] <- future$inventory.high[i]*(as.numeric(input$urbanization)/100)
    future$urbanization.low[i] <- future$inventory.low[i]*(as.numeric(input$urbanization)/100)
    
  }
  
  future$inventory <- round(as.numeric(future$inventory),0)
  future$inventory.high <- round(as.numeric(future$inventory.high),0)
  future$inventory.low <- round(as.numeric(future$inventory.low),0)
  
  future_plt <- select(future, year, inventory, inventory.high, inventory.low)
  
  future_plt <- tidyr::gather(future_plt, 
                              key = 'type', 
                              value = 'forecast', 
                              -year)
  
  future_plt$forecast_div <- future_plt$forecast/1000000000
  future_plt$bounds <- ifelse(future_plt$type == "inventory", "n", "y")
  
  future_plt$bounds <- factor(future_plt$bounds, levels = c("y", "n"))
  
  plt <- ggplot(data = future_plt, aes(x = year, y = forecast_div, linetype = bounds, colour = bounds, size = bounds, group = type)) + 
    geom_line(aes(text = paste0("Year: ", year, "<br>Forecast: ", formattable::comma(forecast)))) +
    scale_colour_manual(values = c("maroon", "darkgreen")) +
    scale_linetype_manual(values = c(3,1)) +
    scale_size_manual(values = c(1,1)) +
    theme(panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          legend.position = 'none') +
    ylab("Inventory (Billions, Cubic Feet)") +
    xlab("") +
    scale_y_continuous(limits = c(0,110))
  
  
  ggplotly(plt, tooltip = 'text')
  
})
```


Row 
-----------------------------------------------------------------------

### Current Inventory (2014) by FIA Unit

```{r echo = F,warning=FALSE,message=FALSE,error=FALSE}
inventory.map <- inv.state.rolled %>%
  select(STATE_ABBR, EVAL_GRP_YEAR, SELECT_WHITE_OAK_INV_CUFT_GS) %>%
  group_by(STATE_ABBR, EVAL_GRP_YEAR) %>%
  summarise(tot = sum(SELECT_WHITE_OAK_INV_CUFT_GS, na.rm = TRUE)) %>%
  filter(EVAL_GRP_YEAR == 2014)

inventory.map.unit <- inv.unit.rolled %>%
  select(STATE_ABBR, EVAL_GRP_YEAR, UNIT, SELECT_WHITE_OAK_INV_CUFT_GS) %>%
  group_by(STATE_ABBR, EVAL_GRP_YEAR, UNIT) %>%
  summarise(tot = sum(SELECT_WHITE_OAK_INV_CUFT_GS)) %>%
  filter(EVAL_GRP_YEAR == 2014)

colnames(inventory.map)[1] <- "STUSPS"

states_shp <- readOGR("MapFiles", layer = "cb_2016_us_state_500k",
                      GDAL1_integer64_policy = TRUE, verbose = FALSE)

states.shp <- subset(states_shp, !states_shp$STUSPS %in% c("AK", "GU", "MP", "VI", "AS", "PR", "HI"))

states.shp <- sp::merge(states.shp, inventory.map, by = "STUSPS")

stusps <- as.vector(as.character(states.shp@data$STUSPS))

units_shp <- readOGR("MapFiles", layer = "eastunits",
                     GDAL1_integer64_policy = TRUE, verbose = FALSE)

units.shp <- sp::merge(units_shp, inventory.map.unit, by = "UNIT")
units.shp@data <- units.shp@data[complete.cases(units.shp@data),]
units.shp@data$tot[units.shp@data$tot == 0] <- min(units.shp@data$tot[units.shp@data$tot > 0])-5000
units.shp@data$tot.log <- log(units.shp@data$tot)

output$inventory_map <- renderLeaflet({
  
  id <- as.vector(states.shp@data$STUSPS)
  
  leaflet(units.shp, options = leafletOptions(zoomControl = FALSE))%>%
    addPolygons(color = '#444444', weight = 1, smoothFactor = 0.5,
                opacity = 1.0, fillOpacity = 0.5,
                fillColor = ~colorNumeric("YlGn", tot.log)(tot.log)) %>%
    addPolygons(data = states.shp, layerId = id, color = '#000000', weight = 1, smoothFactor = 0.5,
                opacity = 1.0, fillOpacity = 0, label = id,
                highlightOptions = highlightOptions(color = "white", weight = 2,
                                                    bringToFront = TRUE))
  
})

leafletOutput('inventory_map')
```


### Inventory Forecast by State *with 95% Confidence Interval*

```{r echo = F,warning=FALSE,message=FALSE,error=FALSE}
renderPlotly({
  
  st <- as.character(input$inventory_map_shape_click$id)
  
  validate(need(try(state.data <-
                      filter(data, STATE_ABBR == st))
                , "Click on a state to the left to view the state-level forecast."))
  
  state.data <- state.data[, c(1, 2, 5:7)] %>%
    group_by(STATE_ABBR, EVAL_GRP_YEAR) %>%
    summarise_all(funs(sum(., na.rm = TRUE)))
  
  state.inv <- inv.state.rolled %>%
    select(STATE_ABBR, EVAL_GRP_YEAR, SELECT_WHITE_OAK_INV_CUFT_GS) %>%
    group_by(STATE_ABBR, EVAL_GRP_YEAR) %>%
    summarise(tot = sum(SELECT_WHITE_OAK_INV_CUFT_GS, na.rm = TRUE))
  
  state.mod <-
    merge(state.data, state.inv, by = c("STATE_ABBR", "EVAL_GRP_YEAR"))
  
  colnames(state.mod)[3:6] <-
    c("total.growth",
      "total.mortality",
      "total.removal",
      "inv.gs")
  
  state.mod$growth.change <- state.mod$total.growth / state.mod$inv.gs
  state.mod$mortality.change <- state.mod$total.mortality / state.mod$inv.gs
  state.mod$removal.change <- state.mod$total.removal / state.mod$inv.gs
  
  state.mod <- filter(state.mod, EVAL_GRP_YEAR <= 2014 & EVAL_GRP_YEAR >= 1991)
  
  #check for consecutive years
  state.mod$conesc <- 2
  
  for(i in 1:(nrow(state.mod)-1)) {
    if((state.mod$EVAL_GRP_YEAR[i+1] == (state.mod$EVAL_GRP_YEAR[i]) + 1)) {
      state.mod$consec[i] <- 0
    } else {
      state.mod$consec[i] <- 1
    }
  }
  
  state.mod$consec[nrow(state.mod)] <- 0 
  
  state.mod <- filter(state.mod, consec == 0)
  
  # Build model
  num <- nrow(state.mod)
  
  validate(need(try(model_growth_st <- lm(growth.change ~ c(1:num), data = state.mod))
                , "There is no forecast available for selected state."))
  model_mortality_st <- lm(mortality.change ~ c(1:num), data = state.mod)
  model_removal_st <- lm(removal.change ~ c(1:num), data = state.mod)
  
  # Store coefficients
  growth_slope <- coef(model_growth_st)[[2]]
  mortality_slope <- coef(model_mortality_st)[[2]]
  removal_slope <- coef(model_removal_st)[[2]]
  
  # Store upper and lower bounds
  growth_slope_upper <- confint(model_growth_st)[4]
  growth_slope_lower <- confint(model_growth_st)[2]
  
  mortality_slope_upper <- confint(model_mortality_st)[4]
  mortality_slope_lower <- confint(model_mortality_st)[2]
  
  removal_slope_upper <- confint(model_removal_st)[4]
  removal_slope_lower <- confint(model_removal_st)[2]
  
  # Use average of 2010 - 2014 as intercept
  for_int <- filter(state.mod, EVAL_GRP_YEAR >= 2010)
  growth_int <- mean(for_int$growth.change)
  mortality_int <- mean(for_int$mortality.change)
  removal_int <- mean(for_int$removal.change)
  
  # Build "forecast" dataframe
  future_st <- NULL
  future_st$year <- 2014:2040
  future_st <- data.frame(future_st)
  
  future_st$inventory <- 0
  future_st$inventory.high <- 0
  future_st$inventory.low <- 0
  future_st$growth <- 0
  future_st$growth.high <- 0
  future_st$growth.low <- 0
  future_st$removals <- 0
  future_st$mortality <- 0
  future_st$mortality.high <- 0
  future_st$mortality.low <- 0
  future_st$urbanization <- 0
  future_st$urbanization.high <- 0
  future_st$urbanization.low <- 0
  
  # Fill in data for 2014
  inv_14 <- state.mod %>%
    filter(EVAL_GRP_YEAR == 2014) %>%
    select(inv.gs)
  
  future_st$inventory[1] <- as.numeric(inv_14[1, 1])
  future_st$inventory.high[1] <- as.numeric(inv_14[1, 1])
  future_st$inventory.low[1] <- as.numeric(inv_14[1, 1])
  future_st$growth[1] <- state.mod$total.growth[nrow(state.mod)]
  future_st$removals[1] <- state.mod$total.removal[nrow(state.mod)]
  future_st$mortality[1] <- state.mod$total.mortality[nrow(state.mod)]
  future_st$growth.high[1] <- state.mod$total.growth[nrow(state.mod)]
  future_st$mortality.high[1] <- state.mod$total.mortality[nrow(state.mod)]
  future_st$growth.low[1] <- state.mod$total.growth[nrow(state.mod)]
  future_st$mortality.low[1] <- state.mod$total.mortality[nrow(state.mod)]
  
  # Calculate growth and mortality % "forecast"
  future_st$dummy <- 0:26
  
  future_st$new_growth <- ((future_st$dummy * growth_slope) + growth_int)
  future_st$new_mortality <- ((future_st$dummy * mortality_slope) + mortality_int)
  
  future_st$high_growth <- ((future_st$dummy * growth_slope_upper) + growth_int)
  future_st$high_mortality <- ((future_st$dummy * mortality_slope_upper) + mortality_int)
  
  future_st$low_growth <- ((future_st$dummy * growth_slope_lower) + growth_int)
  future_st$low_mortality <- ((future_st$dummy * mortality_slope_lower) + mortality_int)
  
  # Cannot be negative % of inventory
  future_st$low_growth <- ifelse(future_st$low_growth <= 0, 0, future_st$low_growth)
  future_st$low_mortality <- ifelse(future_st$low_mortality <= 0, 0, future_st$low_mortality)
  future_st$high_growth <- ifelse(future_st$high_growth <= 0, 0, future_st$high_growth)
  future_st$high_mortality <- ifelse(future_st$high_mortality <= 0, 0, future_st$high_mortality)
  future_st$new_growth <- ifelse(future_st$new_growth <= 0, 0, future_st$new_growth)
  future_st$new_mortality <- ifelse(future_st$new_mortality <= 0, 0, future_st$new_mortality)
  
  future_st$dummy <- NULL
  
  # fill in forecast table row by row
  for (i in 2:nrow(future_st)) {
    future_st$inventory[i] <- future_st$inventory[i - 1] + future_st$growth[i - 1] - future_st$removals[i - 1] - future_st$mortality[i - 1] - future_st$urbanization[i - 1]
    
    future_st$inventory.high[i] <- future_st$inventory.high[i - 1] + future_st$growth.high[i - 1] - future_st$removals[i - 1] - future_st$mortality.low[i - 1] - future_st$urbanization.high[i - 1]
    
    future_st$inventory.low[i] <- future_st$inventory.low[i - 1] + future_st$growth.low[i - 1] - future_st$removals[i - 1] - future_st$mortality.high[i - 1] - future_st$urbanization.low[i - 1]
    
    future_st$growth[i] <- future_st$inventory[i] * (future_st$new_growth[i])
    future_st$growth.high[i] <- future_st$inventory.high[i] * (future_st$high_growth[i])
    future_st$growth.low[i] <- future_st$inventory.low[i] * (future_st$low_growth[i])
    
    future_st$mortality[i] <- future_st$inventory[i] * (future_st$new_mortality[i])
    future_st$mortality.high[i] <- future_st$inventory.low[i] * (future_st$high_mortality[i])
    future_st$mortality.low[i] <- future_st$inventory.high[i] * (future_st$low_mortality[i])
    
    future_st$removals[i] <- future_st$removals[i - 1] * ((1 + as.numeric(input$removal) / 100))
    
    future_st$urbanization[i] <- future_st$inventory[i] * (as.numeric(input$urbanization) / 100)
    future_st$urbanization.high[i] <- future_st$inventory.high[i] * (as.numeric(input$urbanization) / 100)
    future_st$urbanization.low[i] <- future_st$inventory.low[i] * (as.numeric(input$urbanization) / 100)
    
  }
  
  future_st$inventory <- round(as.numeric(future_st$inventory),0)
  future_st$inventory.high <- round(as.numeric(future_st$inventory.high),0)
  future_st$inventory.low <- round(as.numeric(future_st$inventory.low),0)  
  
  future_plt <- select(future_st, year, inventory, inventory.high, inventory.low)
  
  future_plt <- tidyr::gather(future_plt, 
                              key = 'type', 
                              value = 'forecast', 
                              -year)
  
  future_plt$forecast_div <- future_plt$forecast/1000000000
  
  future_plt$bounds <- ifelse(future_plt$type == "inventory", "n", "y")
  
  future_plt$bounds <- factor(future_plt$bounds, levels = c("y", "n"))
  
  plt_st <- ggplot(data = future_plt, aes(x = year, y = forecast_div, linetype = bounds, colour = bounds, size = bounds, group = type)) + 
    geom_line(aes(text = paste0("Year: ", year, "<br>Forecast: ", formattable::comma(forecast)))) +
    scale_colour_manual(values = c("maroon", "darkgreen")) +
    scale_linetype_manual(values = c(3,1)) +
    scale_size_manual(values = c(1,1)) +
    theme(panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          legend.position = 'none') +
    ylab(paste0(unique(state.mod$STATE_ABBR)," Inventory (Billions, Cubic Feet)")) +
    xlab("")
  
  ggplotly(plt_st, tooltip = 'text')
  
})
```


Forecast Data
=======================================================================

Inputs {.sidebar}
-----------------------------------------------------------------------

\

The data displayed to the right includes the values selected for Annual increase in Removal and/or Urbanization from the prior tabs. 

\


```{r}
selectInput("forecast_type", label = "Please select which forecast you would like to view:", choices = c("Forecast", "Alternate Forecast"), selected = "Forecast")
```

Enter Brown-Forman's demand for White Oak in the year `r max(data$EVAL_GRP_YEAR)-3`:

\

```{r}
numericInput("demand", label = paste0("Brown-Forman ", max(data$EVAL_GRP_YEAR)-3," Demand:"), 360000, min = 0, max = 1000000000)
```

\

Use the sliding bar below to adjust the increase in Annual Demand to analyze the portion of Total Inventory and Total Removal that B-F Demand occupies.

```{r}
tags$style(HTML(".js-irs-3 .irs-single, .js-irs-3 .irs-bar-edge, .js-irs-3 .irs-bar {background: green}"))

sliderInput("bf_demand", label = "Brown-Forman Annual Demand Increase (%):",
            min = 0.0, max = 20.0, value = 10.0, step = 0.5)
```

Row
-----------------------------------------------------------------------

```{r echo = F,warning=FALSE,message=FALSE,error=FALSE}
renderTable({
  
  if (input$forecast_type == "Alternate Forecast") {  
    
    future$demand[1] <- as.numeric(input$demand)
    future$`demand % of inventory` <- (future$demand[1]/future$inventory[1])*100
    future$`demand % of removals` <- (future$demand[1]/future$removals[1])*100
    
    # fill in forecast table row by row
    for (i in 2:nrow(future)) {
      future$inventory[i] <- future$inventory[i-1] + future$growth[i-1] - future$removals[i-1] - future$mortality[i-1] - future$urbanization[i-1]
      
      future$inventory.high[i] <- future$inventory.high[i-1] + future$growth.high[i-1] - future$removals[i-1] - future$mortality.low[i-1] - future$urbanization.high[i-1]
      
      future$inventory.low[i] <- future$inventory.low[i-1] + future$growth.low[i-1] - future$removals[i-1] - future$mortality.high[i-1] - future$urbanization.low[i-1]
      
      future$growth[i] <- future$inventory[i]*(future$new_growth[i])
      future$growth.high[i] <- future$inventory.high[i]*(future$high_growth[i])
      future$growth.low[i] <- future$inventory.low[i]*(future$low_growth[i])
      
      future$mortality[i] <- future$inventory[i]*(future$new_mortality[i])
      future$mortality.high[i] <- future$inventory.low[i]*(future$high_mortality[i])
      future$mortality.low[i] <- future$inventory.high[i]*(future$low_mortality[i])
      
      future$removals[i] <- future$removals[i-1]*((1 + as.numeric(input$removal)/100))
      
      future$urbanization[i] <- future$inventory[i]*(as.numeric(input$urbanization)/100)
      future$urbanization.high[i] <- future$inventory.high[i]*(as.numeric(input$urbanization)/100)
      future$urbanization.low[i] <- future$inventory.low[i]*(as.numeric(input$urbanization)/100)
      
      future$demand[i] <- future$demand[i-1]*(1+(as.numeric(input$bf_demand)/100))
      future$`demand % of inventory`[i] <- (future$demand[i]/future$inventory[i])*100
      future$`demand % of removals`[i] <- (future$demand[i]/future$removals[i])*100
      
    }
    
    future <- select(future, year, inventory, growth, removals, mortality, demand, `demand % of inventory`, `demand % of removals`)
    future$inventory <- round(formattable::comma(as.numeric(future$inventory),0))
    future$growth <- round(formattable::comma(as.numeric(future$growth),0))
    future$removals <- round(formattable::comma(as.numeric(future$removals),0))
    future$mortality <- round(formattable::comma(as.numeric(future$mortality),0))
    future$demand <- round(formattable::comma(as.numeric(future$demand),0))
    future$`demand % of inventory` <- formattable::percent(future$`demand % of inventory`, digits = 3)
    future$`demand % of removals` <- formattable::percent(future$`demand % of removals`, digits = 3)
    colnames(future) <- c("Year", "Inventory", "Growth", "Removals", "Mortality", "Demand", "Demand % of Inventory", "Demand % of Removals")
    future
    
  } else {
    
    future_alt$demand[1] <- as.numeric(input$demand)
    future_alt$`demand % of inventory` <- (future_alt$demand[1]/future_alt$inventory[1])*100
    future_alt$`demand % of removals` <- (future_alt$demand[1]/future_alt$removals[1])*100
    
    # fill in forecast table row by row
    for (i in 2:nrow(future_alt)) {
      # Inventory = Inventory + Growth - Removal - Mortality - Urbanization  
      future_alt$inventory[i] <- future_alt$inventory[i-1] + future_alt$growth[i-1] - future_alt$removals[i-1] - future_alt$mortality[i-1] - future_alt$urbanization[i-1]
      
      # Upper Bound Inventory = Upper Bound Inventory + High Growth - Low Removal - Low Mortality - High Urbanization (i.e. the Urbanization calculated from the High Inventory)
      future_alt$inventory.high[i] <- future_alt$inventory.high[i-1] + future_alt$growth.high[i-1] - future_alt$removal.low[i-1] - future_alt$mortality.low[i-1] - future_alt$urbanization.high[i-1]
      
      # Lower Bound Inventory = Lower Bound Inventory + Low Growth - High Removal - High Mortality - Low Urbanization (i.e. the Urbanization calculated from the Low Inventory)
      future_alt$inventory.low[i] <- future_alt$inventory.low[i-1] + future_alt$growth.low[i-1] - future_alt$removal.high[i-1] - future_alt$mortality.high[i-1] - future_alt$urbanization.low[i-1]
      
      future_alt$growth[i] <- future_alt$inventory[i]*(future_alt$new_growth[i])    
      future_alt$growth.high[i] <- future_alt$inventory.high[i]*(future_alt$high_growth[i])
      future_alt$growth.low[i] <- future_alt$inventory.low[i]*(future_alt$low_growth[i])
      
      future_alt$mortality[i] <- future_alt$inventory[i]*(future_alt$new_mortality[i])
      future_alt$mortality.high[i] <- future_alt$inventory.low[i]*(future_alt$high_mortality[i])
      future_alt$mortality.low[i] <- future_alt$inventory.high[i]*(future_alt$low_mortality[i])
      
      future_alt$removals[i] <- future_alt$inventory[i]*future_alt$new_removal[i]
      future_alt$removal.high[i] <- future_alt$inventory.low[i]*future_alt$high_removal[i]
      future_alt$removal.low[i] <- future_alt$inventory.high[i]*future_alt$low_removal[i]
      
      future_alt$urbanization[i] <- future_alt$inventory[i]*(as.numeric(input$urbanization2)/100)
      future_alt$urbanization.high[i] <- future_alt$inventory.high[i]*(as.numeric(input$urbanization2)/100)
      future_alt$urbanization.low[i] <- future_alt$inventory.low[i]*(as.numeric(input$urbanization2)/100)
      
      future_alt$demand[i] <- future_alt$demand[i-1]*(1+(as.numeric(input$bf_demand)/100))
      future_alt$`demand % of inventory`[i] <- (future_alt$demand[i]/future_alt$inventory[i])*100
      future_alt$`demand % of removals`[i] <- (future_alt$demand[i]/future_alt$removals[i])*100
    }
    
    future_alt <- select(future_alt, year, inventory, growth, removals, mortality, demand, `demand % of inventory`, `demand % of removals`)
    future_alt$inventory <- round(formattable::comma(as.numeric(future_alt$inventory),0))
    future_alt$growth <- round(formattable::comma(as.numeric(future_alt$growth),0))
    future_alt$removals <- round(formattable::comma(as.numeric(future_alt$removals),0))
    future_alt$mortality <- round(formattable::comma(as.numeric(future_alt$mortality),0))
    future_alt$demand <- round(formattable::comma(as.numeric(future_alt$demand),0))
    future_alt$`demand % of inventory` <- formattable::percent(future_alt$`demand % of inventory`, digits = 3)
    future_alt$`demand % of removals` <- formattable::percent(future_alt$`demand % of removals`, digits = 3)
    colnames(future_alt) <- c("Year", "Inventory", "Growth", "Removals", "Mortality", "Demand", "Demand % of Inventory", "Demand % of Removals")
    future_alt
    
  }
  
})
```

Simulation
=======================================================================
```{r}
library(tidyverse)
library(ggplot2)
library(readr)
library(tibbletime)
library(corrplot)
library(utf8)
library(stringr)
library(lmtest)
library(car)
library(gvlma)
library(sandwich)
library(forecast)
```

```{r echo = FALSE}
inv23 <- read_csv("WhiteOakData/white_oak_inv_23.csv")
inv24 <- read_csv("WhiteOakData/white_oak_inv_24.csv")
inv33 <- read_csv("WhiteOakData/white_oak_inv_33.csv")

inventory <- rbind(inv23, inv24, inv33)

inventory[is.na(inventory)] <- 0

inventory$UNIT <- paste0(inventory$STATECD, "0", inventory$UNITCD)

oak <- read_csv("WhiteOakData/white-oak-clean.csv")
oak[is.na(oak)] <- 0

oak_merge <- oak %>%
  group_by(STATE_ABBR, EVAL_GRP_YEAR, UNIT) %>%
  summarise(total_removal = sum(SELECT_WHITE_OAK_REMOVAL, na.rm = TRUE),
            total_growth = sum(SELECT_WHITE_OAK_GROWTH, na.rm = TRUE),
            total_mortality = sum(SELECT_WHITE_OAK_MORTALITY, na.rm = TRUE)) %>%
  mutate(total_growth_true = total_growth + total_mortality)

oak_merge$total_growth <- NULL
colnames(oak_merge)[6] <- "total_growth"

inventory_merge <- inventory %>%
  group_by(STATE_ABBR, EVAL_GRP_YEAR, UNIT) %>%
  summarise(total_inventory = sum(SELECT_WHITE_OAK_INV_CUFT_GS, na.rm = TRUE),
            total_hardwood = sum(TOTAL_HARDWOOD_INV_CUFT_GS, na.rm = TRUE))

oak_total <- merge(oak_merge, inventory_merge, by = c("STATE_ABBR", "EVAL_GRP_YEAR", "UNIT"))

fire <- read_csv("fire_data.csv")
colnames(fire) <- c("STATE_ABBR", "num_fires", "acres_burned", "EVAL_GRP_YEAR")

mod.dat <- oak_total %>%
  group_by(STATE_ABBR, EVAL_GRP_YEAR) %>%
  summarise(total_removal = sum(total_removal),
            total_growth = sum(total_growth),
            total_mortality = sum(total_mortality),
            total_inventory = sum(total_inventory),
            total_hardwood = sum(total_hardwood))

r9 <- c("MN", "IA", "MO", "WI", "IL", "MI", "IN", "OH", "WV", "PA", "NJ", "DE", "MD", "NY", "CT", "RI", "MA", "VT", "NH", "ME")

r2 <- c("ND", "SD", "NE", "KS")

r8 <- c("TX", "OK", "LA", "AR", "MS", "AL", "GA", "FL", "TN", "KY", "VA", "NC", "SC")

mod.dat$region <- ifelse(mod.dat$STATE_ABBR %in% r9, "R9",
                         ifelse(mod.dat$STATE_ABBR %in% r8, "R8",
                                ifelse(mod.dat$STATE_ABBR %in% r2, "R2", "NA")))

write_csv(mod.dat, "mod_dat_final.csv")


pdsi_states <- read_csv("pdsi_states.csv")

pest_output <- read_csv("pest_output.csv")
pest_merge <- merge(pest_output, pdsi_states, by = "State")
colnames(pest_merge)[5] <- "STATE_ABBR"
pest_merge$region <- ifelse(pest_merge$STATE_ABBR %in% r9, "R9",
                         ifelse(pest_merge$STATE_ABBR %in% r8, "R8",
                                ifelse(pest_merge$STATE_ABBR %in% r2, "R2", "NA")))

state_land_area <- read_csv("state_land_area.csv")

state_land_area <- separate(data = state_land_area, col = `State`, into = c("State", "Other"), sep = " flag")
state_land_area <- separate(data = state_land_area, col = `Land Area (sq mi)`, into = c("land_area", "other"), sep = "\\.")
state_land_area <- state_land_area %>%
  select(State, land_area)

state_land_area$land_area <- gsub(",", "", state_land_area$land_area)
state_land_area$land_area2 <- as.numeric(state_land_area$land_area)
state_land_area$land_area <- state_land_area$land_area2
state_land_area$land_area2 <- NULL

state_land_area_merge <- merge(state_land_area, pdsi_states, by = c("State")) %>%
  mutate(land_area_acres = land_area*640)

state_land_area_merge$land_area <- NULL
state_land_area_merge$State <- NULL

colnames(state_land_area_merge) <- c("STATE_ABBR", "land_area_acres")

r9 <- c("MN", "IA", "MO", "WI", "IL", "MI", "IN", "OH", "WV", "PA", "NJ", "DE", "MD", "NY", "CT", "RI", "MA", "VT", "NH", "ME")

r2 <- c("ND", "SD", "NE", "KS")

r8 <- c("TX", "OK", "LA", "AR", "MS", "AL", "GA", "FL", "TN", "KY", "VA", "NC", "SC")

state_land_area_merge$region <- ifelse(state_land_area_merge$STATE_ABBR %in% r9, "R9",
                         ifelse(state_land_area_merge$STATE_ABBR %in% r8, "R8",
                                ifelse(state_land_area_merge$STATE_ABBR %in% r2, "R2", "NA")))

state_land_area_reg <- state_land_area_merge %>%
  group_by(region) %>%
  summarise(land_area_acres = sum(land_area_acres))

```

```{r echo = FALSE}

# catastrophic_fires <- read_csv("catastrophic_fires.csv")
# 
# catastrophic_fires <- mutate(catastrophic_fires,
#                              cu_ft_burned = acres_burned*43560)
# 
# state_land_area <- read_csv("state_land_area.csv")
# 
# state_land_area <- separate(data = state_land_area, col = `State`, into = c("State", "Other"), sep = " flag")
# state_land_area <- separate(data = state_land_area, col = `Land Area (sq mi)`, into = c("land_area", "other"), sep = "\\.")
# state_land_area <- state_land_area %>%
#   select(State, land_area)
# 
# state_land_area$land_area <- gsub(",", "", state_land_area$land_area)
# state_land_area$land_area2 <- as.numeric(state_land_area$land_area)
# state_land_area$land_area <- state_land_area$land_area2
# state_land_area$land_area2 <- NULL
# 

# 
# state_land_area_merge <- merge(state_land_area, pdsi_states, by = c("State")) %>%
#   mutate(land_area_acres = land_area*640)
# 
# state_land_area_merge$land_area <- NULL
# state_land_area_merge$State <- NULL
# 
# colnames(state_land_area_merge) <- c("STATE_ABBR", "land_area_acres")
# 
# r9 <- c("MN", "IA", "MO", "WI", "IL", "MI", "IN", "OH", "WV", "PA", "NJ", "DE", "MD", "NY", "CT", "RI", "MA", "VT", "NH", "ME")
# 
# r2 <- c("ND", "SD", "NE", "KS")
# 
# r8 <- c("TX", "OK", "LA", "AR", "MS", "AL", "GA", "FL", "TN", "KY", "VA", "NC", "SC")
# 
# state_land_area_merge$region <- ifelse(state_land_area_merge$STATE_ABBR %in% r9, "R9",
#                          ifelse(state_land_area_merge$STATE_ABBR %in% r8, "R8",
#                                 ifelse(state_land_area_merge$STATE_ABBR %in% r2, "R2", "NA")))
# 
# catastrophic_fires$region <- ifelse(catastrophic_fires$STATE_ABBR %in% r9, "R9",
#                          ifelse(catastrophic_fires$STATE_ABBR %in% r8, "R8",
#                                 ifelse(catastrophic_fires$STATE_ABBR %in% r2, "R2", "NA")))
# 
# state_land_area_reg <- state_land_area_merge %>%
#   group_by(region) %>%
#   summarise(land_area_acres = sum(land_area_acres))
# 
# catastrophic_fires <- catastrophic_fires %>%
#   filter(region != "NA",
#          EVAL_GRP_YEAR <= 2014) %>%
#   merge(., state_land_area_reg, by = "region")
# 
# catastrophic_fires <- mutate(catastrophic_fires,
#                              pct_area_burned = (acres_burned / land_area_acres))

# mod.dat$region <- ifelse(mod.dat$STATE_ABBR %in% r9, "R9",
#                          ifelse(mod.dat$STATE_ABBR %in% r8, "R8",
#                                 ifelse(mod.dat$STATE_ABBR %in% r2, "R2", "NA")))



pest_merge <- read_csv("pest_final.csv")

pest_r8 <- pest_merge %>% 
  filter(region == "R8") %>% 
  group_by(Year) %>% 
  summarise(total_acres = sum(Acres))

pest_r9 <- pest_merge %>% 
  filter(region == "R9") %>% 
  group_by(Year) %>% 
  summarise(total_acres = sum(Acres))

pest_r2 <- pest_merge %>% 
  filter(region == "R2") %>% 
  group_by(Year) %>% 
  summarise(total_acres = sum(Acres))


# How many times was the pest volume in each region greater than the region's average? (a potential way to quatify the  more rare outbreaks)

pest_r9$outbreak <- ifelse(pest_r9$total_acres > mean(pest_r9$total_acres), 1, 0)
pest_r8$outbreak <- ifelse(pest_r8$total_acres > mean(pest_r8$total_acres), 1, 0)
pest_r2$outbreak <- ifelse(pest_r2$total_acres > mean(pest_r2$total_acres), 1, 0)

pest_inv <- mod.dat %>% 
  filter(EVAL_GRP_YEAR == 2014) %>% 
  group_by(EVAL_GRP_YEAR, region) %>% 
  summarise(total_hardwood = sum(total_hardwood, na.rm = TRUE))

pest_r2$reg_acres <- state_land_area_reg$land_area_acres[state_land_area_reg$region == "R2"]
pest_r9$reg_acres <- state_land_area_reg$land_area_acres[state_land_area_reg$region == "R9"]
pest_r8$reg_acres <- state_land_area_reg$land_area_acres[state_land_area_reg$region == "R8"]

pest_r2$pct_acres <- pest_r2$total_acres/pest_r2$reg_acres
pest_r9$pct_acres <- pest_r9$total_acres/pest_r9$reg_acres
pest_r8$pct_acres <- pest_r8$total_acres/pest_r8$reg_acres

# How often was the state/region in severe drought? 

drought_monitor <- read_csv("drought_monitor.csv")
drought_monitor$year <- str_sub(drought_monitor$ReleaseDate, 1, 4)
colnames(drought_monitor)[2] <- "STATE_ABBR"


drought_monitor$region <- ifelse(drought_monitor$STATE_ABBR %in% r9, "R9",
                         ifelse(drought_monitor$STATE_ABBR %in% r8, "R8",
                                ifelse(drought_monitor$STATE_ABBR %in% r2, "R2", "NA")))

drought_monitor_reg <- drought_monitor %>%
  group_by(region, year) %>%
  filter(region != "NA") %>%
  summarise(total_d4 = sum(as.numeric(D4), na.rm = TRUE),
            total_d3 = sum(as.numeric(D3), na.rm = TRUE))

tornado_dat <- read_csv("Tornado/Consolidated_Tornados_50_16.csv")

colnames(tornado_dat)[2] <- "STATE_ABBR"
tornado_dat$region <- ifelse(tornado_dat$STATE_ABBR %in% r9, "R9",
                         ifelse(tornado_dat$STATE_ABBR %in% r8, "R8",
                                ifelse(tornado_dat$STATE_ABBR %in% r2, "R2", "NA")))

tornado_dat_5 <- tornado_dat %>% 
  filter(magnitude == 5) %>% 
  group_by(region, year) %>% 
  summarise(acres_destroyed = sum(acres))

tornado_2 <- tornado_dat_5 %>% 
  filter(region == "R2") %>% 
  mutate(region_acres = state_land_area_reg$land_area_acres[state_land_area_reg$region == "R2"],
         pct_acres = acres_destroyed/region_acres)

tornado_9 <- tornado_dat_5 %>% 
  filter(region == "R9") %>% 
  mutate(region_acres = state_land_area_reg$land_area_acres[state_land_area_reg$region == "R9"],
         pct_acres = acres_destroyed/region_acres)

tornado_8 <- tornado_dat_5 %>% 
  filter(region == "R8") %>% 
  mutate(region_acres = state_land_area_reg$land_area_acres[state_land_area_reg$region == "R8"],
         pct_acres = acres_destroyed/region_acres)

#mod.dat <- read_csv("mod_dat_final.csv")
#catastrophic_fires <- read_csv("cat_fires_final.csv")

more_fire <- read_csv("more_fire_final.csv")

fire_inv <- mod.dat  %>% 
  filter(EVAL_GRP_YEAR == 2014) %>% 
  group_by(EVAL_GRP_YEAR, region) %>% 
  summarise(total_inv = sum(total_inventory, na.rm = TRUE))

fire_r8_df <- more_fire %>% 
  filter(region == "R8") %>% 
  group_by(FIRE_YEAR) %>% 
  summarise(total_acres = sum(tot_acres))

fire_r9_df <- more_fire %>% 
  filter(region == "R9") %>% 
  group_by(FIRE_YEAR) %>% 
  summarise(total_acres = sum(tot_acres))

fire_r2_df <- more_fire %>% 
  filter(region == "R2") %>% 
  group_by(FIRE_YEAR) %>% 
  summarise(total_acres = sum(tot_acres))

fire_r2_df$reg_acres <- fire_inv$total_inv[fire_inv$region == "R2"]
fire_r9_df$reg_acres <- fire_inv$total_inv[fire_inv$region == "R9"]
fire_r8_df$reg_acres <- fire_inv$total_inv[fire_inv$region == "R8"]

fire_r2_df$pct_acres <- fire_r2_df$total_acres/fire_r2_df$reg_acres
fire_r9_df$pct_acres <- fire_r9_df$total_acres/fire_r9_df$reg_acres
fire_r8_df$pct_acres <- fire_r8_df$total_acres/fire_r8_df$reg_acres

## Urbanized land 2050

pct_urban <- read_csv("percent-forestland-urban.csv")

# weight by state percent of WO in region

region_weights <- mod.dat %>% 
  filter(EVAL_GRP_YEAR == 2014) %>% 
  group_by(STATE_ABBR, region) %>% 
  summarise(total_inventory = sum(total_inventory))

region_weights$region_total <- ifelse(region_weights$region == "R9", sum(region_weights$total_inventory[region_weights$region == "R9"]), ifelse(region_weights$region == "R8", sum(region_weights$total_inventory[region_weights$region == "R8"]), sum(region_weights$total_inventory[region_weights$region == "R2"])))

region_weights$WO_pct_region <- region_weights$total_inventory/region_weights$region_total

colnames(pct_urban)[1] <- "STATE_ABBR"

pct_urban_final <- merge(pct_urban, region_weights, by = "STATE_ABBR") %>% 
  mutate(mult_weight = pct_forestland_urban_2050 * WO_pct_region) %>% 
  group_by(region) %>% 
  summarise(weight_sum = sum(WO_pct_region),
            weighted_pct = sum(mult_weight)) %>% 
  mutate(weighted_avg = weighted_pct / weight_sum)




mod.dat.r9 <- mod.dat %>% 
  filter(region == "R9") %>% 
  group_by(EVAL_GRP_YEAR) %>% 
  summarise(total_inventory = sum(total_inventory),
            total_mortality = sum(total_mortality),
            total_removal = sum(as.numeric(total_removal)),
            total_growth = sum(as.numeric(total_growth)))

mod.dat.r8 <- mod.dat %>% 
  filter(region == "R8") %>% 
  group_by(EVAL_GRP_YEAR) %>% 
  summarise(total_inventory = sum(total_inventory),
            total_mortality = sum(total_mortality),
            total_removal = sum(as.numeric(total_removal)),
            total_growth = sum(as.numeric(total_growth)))

mod.dat.r2 <- mod.dat %>% 
  filter(region == "R2") %>% 
  group_by(EVAL_GRP_YEAR) %>% 
  summarise(total_inventory = sum(total_inventory),
            total_mortality = sum(total_mortality),
            total_removal = sum(as.numeric(total_removal)),
            total_growth = sum(as.numeric(total_growth)))


```

In the simulation outline below, we sample from a normal distribution with a mean equal to the coefficient of a linear model and a standard deviation equal to the standard error on the coefficient estimate in order to obtain rate of change values the growth, removal, and mortality percent of inventory. We then calculate the new percent of inventory each year by multiplying the number of years elapsed by the rate of chance (ROC) and add an intercept (calculated by taking the mean of the last 5 years of data). We also include a yearly shock value to provide a simple example of how we can incorporate the outside variables like wildfire, pests, drought, etc. The initial inventory value is the true inventory in the year 2014, and each subsequent year is calculated by adding/subtracting the growth, removal, mortality, and yearly shock values.

```{r evaluate = FALSE}

# Calculate growth, removal, mortality percent of inventory
mod.dat.r9 <- mod.dat.r9 %>% 
  mutate(growth.pct = total_growth/total_inventory,
         remv.pct = total_removal/total_inventory, 
         mort.pct = total_mortality/total_inventory)

mod.dat.r8 <- mod.dat.r8 %>% 
  mutate(growth.pct = total_growth/total_inventory,
         remv.pct = total_removal/total_inventory, 
         mort.pct = total_mortality/total_inventory)

mod.dat.r2 <- mod.dat.r2 %>% 
  mutate(growth.pct = total_growth/total_inventory,
         remv.pct = total_removal/total_inventory, 
         mort.pct = total_mortality/total_inventory)

# Filter to 2002 to 2014 for consistency in measurement method

mod.dat.r9 <- filter(mod.dat.r9, EVAL_GRP_YEAR >= 2002 & EVAL_GRP_YEAR <= 2014)
mod.dat.r8 <- filter(mod.dat.r8, EVAL_GRP_YEAR >= 2002 & EVAL_GRP_YEAR <= 2014)
mod.dat.r2 <- filter(mod.dat.r2, EVAL_GRP_YEAR >= 2002 & EVAL_GRP_YEAR <= 2014)

# Build linear models for each and save summary stats
growth.mod.r9 <- lm(growth.pct ~ EVAL_GRP_YEAR, data = mod.dat.r9)
growth.summary.r9 <- summary(growth.mod.r9)
mort.mod.r9 <- lm(mort.pct ~ EVAL_GRP_YEAR, data = mod.dat.r9)
mort.summary.r9 <- summary(mort.mod.r9)
remv.mod.r9 <- lm(remv.pct ~ EVAL_GRP_YEAR, data = mod.dat.r9)
remv.summary.r9 <- summary(remv.mod.r9)

growth.mod.r8 <- lm(growth.pct ~ EVAL_GRP_YEAR, data = mod.dat.r8)
growth.summary.r8 <- summary(growth.mod.r8)
mort.mod.r8 <- lm(mort.pct ~ EVAL_GRP_YEAR, data = mod.dat.r8)
mort.summary.r8 <- summary(mort.mod.r8)
remv.mod.r8 <- lm(remv.pct ~ EVAL_GRP_YEAR, data = mod.dat.r8)
remv.summary.r8 <- summary(remv.mod.r8)

growth.mod.r2 <- lm(growth.pct ~ EVAL_GRP_YEAR, data = mod.dat.r2)
growth.summary.r2 <- summary(growth.mod.r2)
mort.mod.r2 <- lm(mort.pct ~ EVAL_GRP_YEAR, data = mod.dat.r2)
mort.summary.r2 <- summary(mort.mod.r2)
remv.mod.r2 <- lm(remv.pct ~ EVAL_GRP_YEAR, data = mod.dat.r2)
remv.summary.r2 <- summary(remv.mod.r2)

# Check Assumptions

# Independence of Errors (Autocorrelation)

# Breusch Godfrey test
bgtest(remv.mod.r2) # fail to reject
bgtest(remv.mod.r9) # fail to reject
bgtest(remv.mod.r8) # fail to reject
bgtest(mort.mod.r2) # fail to reject
bgtest(mort.mod.r9) # fail to reject
bgtest(mort.mod.r8) # fail to reject
bgtest(growth.mod.r2) # fail to reject
bgtest(growth.mod.r9) # fail to reject
bgtest(growth.mod.r8) # fail to reject

# Ljung Box Test
Box.test(remv.mod.r2$residuals, type="Ljung-Box") # fail to reject
Box.test(remv.mod.r9$residuals, type="Ljung-Box") # fail to reject
Box.test(remv.mod.r8$residuals, type="Ljung-Box") # fail to reject
Box.test(mort.mod.r2$residuals, type="Ljung-Box") # fail to reject
Box.test(mort.mod.r9$residuals, type="Ljung-Box") # fail to reject
Box.test(mort.mod.r8$residuals, type="Ljung-Box") # fail to reject
Box.test(growth.mod.r2$residuals, type="Ljung-Box") # fail to reject
Box.test(growth.mod.r9$residuals, type="Ljung-Box") # fail to reject
Box.test(growth.mod.r8$residuals, type="Ljung-Box") # fail to reject

# Heteroscedasticity

# Breusch Pagan test

bptest(growth.mod.r9) # fail to reject
bptest(growth.mod.r8) # fail to reject
bptest(growth.mod.r2) # fail to reject
bptest(mort.mod.r9) # fail to reject
bptest(mort.mod.r8) # fail to reject
bptest(mort.mod.r2) # fail to reject 
bptest(remv.mod.r9) # fail to reject
bptest(remv.mod.r8) # reject the null
bptest(remv.mod.r2) # reject the null

# Normality
shapiro.test(growth.mod.r9$residuals) # fail to reject
shapiro.test(growth.mod.r8$residuals) # fail to reject
shapiro.test(growth.mod.r2$residuals) # fail to reject
shapiro.test(mort.mod.r9$residuals) # fail to reject
shapiro.test(mort.mod.r8$residuals) # fail to reject
shapiro.test(mort.mod.r2$residuals) # fail to reject 
shapiro.test(remv.mod.r9$residuals) # fail to reject
shapiro.test(remv.mod.r8$residuals) # fail to reject
shapiro.test(remv.mod.r2$residuals) # fail to reject

# Mean of errors is 0
mean(growth.mod.r9$residuals) # true
mean(growth.mod.r8$residuals) # true
mean(growth.mod.r2$residuals) # true
mean(mort.mod.r9$residuals) # true
mean(mort.mod.r8$residuals) # true
mean(mort.mod.r2$residuals) # true
mean(remv.mod.r9$residuals) # true
mean(remv.mod.r8$residuals) # true
mean(remv.mod.r2$residuals) # true

# Check Residuals




# Fix heteroscedasticity problems
remv.mod.r8$newse <- vcovHC(remv.mod.r8, type="HC3")
remv.summary.r8 <- coeftest(remv.mod.r8, remv.mod.r8$newse)
remv.mod.r2$newse <- vcovHC(remv.mod.r2, type="HC3")
remv.summary.r2 <- coeftest(remv.mod.r2, remv.mod.r2$newse)

```

```{r evaluate = FALSE}

years <- 2014:2064

# Build simulation function
run_sim <- function() {
  # Define forecast period
  inv <- 0
  df <- tibble(year = years, inventory.r9 = inv, inventory.r8 = inv, inventory.r2 = inv, total_inventory = inv, new_growth_pct_r9 = inv, new_growth_pct_r8 = inv, new_remv_pct_r8 = inv, new_mort_pct_r2 = inv, new_mort_pct_r8 = inv)
  
  # Initial inventory equal to 2014 inventory
  df$inventory.r9[df$year == 2014] <- mod.dat.r9$total_inventory[nrow(mod.dat.r9)]
  df$inventory.r8[df$year == 2014] <- mod.dat.r8$total_inventory[nrow(mod.dat.r8)]
  df$inventory.r2[df$year == 2014] <- mod.dat.r2$total_inventory[nrow(mod.dat.r2)]
  
  df$new_growth_pct_r8[df$year == 2014] <- mod.dat.r8$growth.pct[nrow(mod.dat.r8)]
  df$new_growth_pct_r9[df$year == 2014] <- mod.dat.r9$growth.pct[nrow(mod.dat.r9)]
  df$new_mort_pct_r8[df$year == 2014] <- mod.dat.r8$mort.pct[nrow(mod.dat.r8)]
  df$new_mort_pct_r2[df$year == 2014] <- mod.dat.r2$mort.pct[nrow(mod.dat.r2)]
  df$new_remv_pct_r8[df$year == 2014] <- mod.dat.r8$remv.pct[nrow(mod.dat.r8)]
  
  # Build the normal distributions for growth, mortality, removal
  growth_dist_r9 <- rnorm(100, growth.mod.r9$coefficients[[2]], growth.summary.r9$coefficients[[4]])
  mortality_dist_r9 <- rnorm(100, mort.mod.r9$coefficients[[2]], mort.summary.r9$coefficients[[4]])
  removal_dist_r9 <- rnorm(100, remv.mod.r9$coefficients[[2]], remv.summary.r9$coefficients[[4]])
  
  growth_dist_r8 <- rnorm(100, growth.mod.r8$coefficients[[2]], growth.summary.r8$coefficients[[4]])
  mortality_dist_r8 <- rnorm(100, mort.mod.r8$coefficients[[2]], mort.summary.r8$coefficients[[4]])
  removal_dist_r8 <- rnorm(100, remv.mod.r8$coefficients[[2]], remv.summary.r8[[4]])

  growth_dist_r2 <- rnorm(100, growth.mod.r2$coefficients[[2]], growth.summary.r2$coefficients[[4]])
  mortality_dist_r2 <- rnorm(100, mort.mod.r2$coefficients[[2]], mort.summary.r2$coefficients[[4]])
  removal_dist_r2 <- rnorm(100, remv.mod.r2$coefficients[[2]], remv.summary.r2[[4]])
  
  # Build additional distributions for the models containing lagged variables
  # removal_lag_dist_r8 <- rnorm(100, remv.mod.r8$coefficients[[3]], remv.summary.r8[[6]])
  # mortality_lag_dist_r8 <- rnorm(100, mort.mod.r8$coefficients[[3]], mort.summary.r8$coefficients[[6]])
  # mortality_lag_dist_r2 <- rnorm(100, mort.mod.r2$coefficients[[3]], mort.summary.r2$coefficients[[6]])
  # growth_lag_dist_r8 <- rnorm(100, growth.mod.r8$coefficients[[3]], growth.summary.r8$coefficients[[6]])
  # growth_lag_dist_r9 <- rnorm(100, growth.mod.r9$coefficients[[3]], growth.summary.r9[[6]])
  
  # Define rate of change by sampling from distributions created above. We want the rate of change to be the same over the entire forecasted time period, but different for each simulation. This is why we define it prior to the while loop below.
  growth_roc_r9 <- base::sample(growth_dist_r9, 1)
  mortality_roc_r9 <- base::sample(mortality_dist_r9, 1)
  removal_roc_r9 <- base::sample(removal_dist_r9, 1)
  
  growth_roc_r8 <- base::sample(growth_dist_r8, 1)
  mortality_roc_r8 <- base::sample(mortality_dist_r8, 1)
  removal_roc_r8 <- base::sample(removal_dist_r8, 1)
  
  growth_roc_r2 <- base::sample(growth_dist_r2, 1)
  mortality_roc_r2 <- base::sample(mortality_dist_r2, 1)
  removal_roc_r2 <- base::sample(removal_dist_r2, 1)
  
  # growth_lag_roc_r8 <- growth.mod.r8$coefficients[[3]]
  # growth_lag_roc_r9 <- growth.mod.r9$coefficients[[3]]
  # mortality_lag_roc_r2 <- mort.mod.r2$coefficients[[3]]
  # mortality_lag_roc_r8 <- mort.mod.r8$coefficients[[3]]
  # removal_lag_roc_r8 <- remv.mod.r8$coefficients[[3]]
  
  # Define rate of change for lagged variables
  # growth_lag_roc_r9 <- base::sample(growth_lag_dist_r9, 1)
  # growth_lag_roc_r8 <- base::sample(growth_lag_dist_r8, 1)
  # mortality_lag_roc_r8 <- base::sample(mortality_lag_dist_r8, 1)
  # mortality_lag_roc_r2 <- base::sample(mortality_lag_dist_r2, 1)
  # removal_lag_roc_r8 <- base::sample(removal_lag_dist_r8, 1)
  
  y <- 2015
  
  # Define intercepts by taking the average of the last 5 years for growth, removal, mortality percent of inventory
  growth_int_r9 <- mean(mod.dat.r9$growth.pct[(nrow(mod.dat.r9)-5):nrow(mod.dat.r9)])
  mortality_int_r9 <- mean(mod.dat.r9$mort.pct[(nrow(mod.dat.r9)-5):nrow(mod.dat.r9)])
  removal_int_r9 <- mean(mod.dat.r9$remv.pct[(nrow(mod.dat.r9)-5):nrow(mod.dat.r9)])
  
  growth_int_r8 <- mean(mod.dat.r8$growth.pct[(nrow(mod.dat.r8)-5):nrow(mod.dat.r8)])
  mortality_int_r8 <- mean(mod.dat.r8$mort.pct[(nrow(mod.dat.r8)-5):nrow(mod.dat.r8)])
  removal_int_r8 <- mean(mod.dat.r8$remv.pct[(nrow(mod.dat.r8)-5):nrow(mod.dat.r8)])
  
  growth_int_r2 <- mean(mod.dat.r2$growth.pct[(nrow(mod.dat.r2)-5):nrow(mod.dat.r2)])
  mortality_int_r2 <- mean(mod.dat.r2$mort.pct[(nrow(mod.dat.r2)-5):nrow(mod.dat.r2)])
  removal_int_r2 <- mean(mod.dat.r2$remv.pct[(nrow(mod.dat.r2)-5):nrow(mod.dat.r2)])
  
  # Determine probability of catastrophic fire in each region
  
  fire_r8_prob <- nrow(fire_r8_df)/(2014-min(fire_r8_df$FIRE_YEAR))
  fire_r9_prob <- nrow(fire_r9_df)/(2014-min(fire_r9_df$FIRE_YEAR))
  fire_r2_prob <- nrow(fire_r2_df)/(2014-min(fire_r2_df$FIRE_YEAR))
  
  fire_density_r9 <- density(fire_r9_df$pct_acres)
  fire_density_r9 <- data.frame(fire_density_r9$x, fire_density_r9$y)
  colnames(fire_density_r9) <- c("x","y")
  fire_density_r9 <- filter(fire_density_r9, x >= 0)
  
  fire_density_r8 <- density(fire_r8_df$pct_acres)
  fire_density_r8 <- data.frame(fire_density_r8$x, fire_density_r8$y)
  colnames(fire_density_r8) <- c("x","y")
  fire_density_r8 <- filter(fire_density_r8, x >= 0)
  
  fire_density_r2 <- density(fire_r2_df$pct_acres)
  fire_density_r2 <- data.frame(fire_density_r2$x, fire_density_r2$y)
  colnames(fire_density_r2) <- c("x","y")
  fire_density_r2 <- filter(fire_density_r2, x >= 0)
  
  # Determine probability of larger than normal pest outbreak
  
  pest_r8_prob <- sum(pest_r8$outbreak)/(max(pest_r8$Year)-min(pest_r8$Year))
  pest_r9_prob <- sum(pest_r9$outbreak)/(max(pest_r9$Year)-min(pest_r9$Year))
  pest_r2_prob <- sum(pest_r2$outbreak)/(max(pest_r2$Year)-min(pest_r2$Year))
  
  pest_density_r2 <- density(pest_r2$pct_acres[pest_r2$outbreak==1])
  pest_density_r2 <- data.frame(pest_density_r2$x, pest_density_r2$y)
  colnames(pest_density_r2) <- c("x","y")
  pest_density_r2 <- filter(pest_density_r2, x >= 0)
  
  pest_density_r8 <- density(pest_r8$pct_acres[pest_r8$outbreak==1])
  pest_density_r8 <- data.frame(pest_density_r8$x, pest_density_r8$y)
  colnames(pest_density_r8) <- c("x","y")
  pest_density_r8 <- filter(pest_density_r8, x >= 0)
  
  pest_density_r9 <- density(pest_r9$pct_acres[pest_r9$outbreak==1])
  pest_density_r9 <- data.frame(pest_density_r9$x, pest_density_r9$y)
  colnames(pest_density_r9) <- c("x","y")
  pest_density_r9 <- filter(pest_density_r9, x >= 0)
  
    # Determine probability of category 5 tornado
  
  tornado_r8_prob <- nrow(tornado_8)/(max(tornado_8$year)-min(tornado_8$year))
  tornado_r9_prob <- nrow(tornado_9)/(max(tornado_9$year)-min(tornado_9$year))
  tornado_r2_prob <- nrow(tornado_2)/(max(tornado_2$year)-min(tornado_2$year))
  
  tornado_density_r2 <- density(tornado_2$pct_acres)
  tornado_density_r2 <- data.frame(tornado_density_r2$x, tornado_density_r2$y)
  colnames(tornado_density_r2) <- c("x","y")
  tornado_density_r2 <- filter(tornado_density_r2, x >= 0)
  
  tornado_density_r8 <- density(tornado_8$pct_acres)
  tornado_density_r8 <- data.frame(tornado_density_r8$x, tornado_density_r8$y)
  colnames(tornado_density_r8) <- c("x","y")
  tornado_density_r8 <- filter(tornado_density_r8, x >= 0)
  
  tornado_density_r9 <- density(tornado_9$pct_acres)
  tornado_density_r9 <- data.frame(tornado_density_r9$x, tornado_density_r9$y)
  colnames(tornado_density_r9) <- c("x","y")
  tornado_density_r9 <- filter(tornado_density_r9, x >= 0)
  
  
  # Use while loop to build forecast year by year
  while (y < 2065) {
    
    # Will there be a fire, and if yes, what will the effect be?
    
    fire_r9 <- sample(c(0,1),1,prob = c((1-fire_r9_prob),fire_r9_prob))
    fire_r8 <- sample(c(0,1),1,prob = c((1-fire_r8_prob),fire_r8_prob))
    fire_r2 <- sample(c(0,1),1,prob = c((1-fire_r2_prob),fire_r2_prob))
    
    if(fire_r9 == 1){
      fire_effect_r9 <- sample(fire_density_r9$x, 1, prob = fire_density_r9$y)
    } else {
      fire_effect_r9 <- 0
    }
    
    if(fire_r8 == 1){
      fire_effect_r8 <- sample(fire_density_r8$x, 1, prob = fire_density_r8$y)
    } else {
      fire_effect_r8 <- 0
    }
    
    if(fire_r2 == 1){
      fire_effect_r2 <- sample(fire_density_r2$x, 1, prob = fire_density_r2$y)
    } else {
      fire_effect_r2 <- 0
    }
    
    # Will there be a pest outbreak? And if yes, what will the effect be?
    
    pest_ob_r9 <- sample(c(0,1),1,prob = c((1-pest_r9_prob),pest_r9_prob))
    pest_ob_r8 <- sample(c(0,1),1,prob = c((1-pest_r8_prob),pest_r8_prob))
    pest_ob_r2 <- sample(c(0,1),1,prob = c((1-pest_r2_prob),pest_r2_prob))
    
    if(pest_ob_r9 == 1){
      pest_effect_r9 <- sample(pest_density_r9$x, 1, prob = pest_density_r9$y)
    } else {
      pest_effect_r9 <- 0
    }
    
    if(pest_ob_r2 == 1){
      pest_effect_r2 <- sample(pest_density_r2$x, 1, prob = pest_density_r2$y)
    } else {
      pest_effect_r2 <- 0
    }
    
    if(pest_ob_r8 == 1){
      pest_effect_r8 <- sample(pest_density_r8$x, 1, prob = pest_density_r8$y)
    } else {
      pest_effect_r8 <- 0
    }
    
    # Will there be a big tornado outbreak? And if yes, what will the effect be?
    
    tornado_r9 <- sample(c(0,1),1,prob = c((1-tornado_r9_prob),tornado_r9_prob))
    tornado_r8 <- sample(c(0,1),1,prob = c((1-tornado_r8_prob),tornado_r8_prob))
    tornado_r2 <- sample(c(0,1),1,prob = c((1-tornado_r2_prob),tornado_r2_prob))
    
    if(tornado_r9 == 1){
      tornado_effect_r9 <- sample(tornado_density_r9$x, 1, prob = tornado_density_r9$y)
    } else {
      tornado_effect_r9 <- 0
    }
    
    if(tornado_r2 == 1){
      tornado_effect_r2 <- sample(tornado_density_r2$x, 1, prob = tornado_density_r2$y)
    } else {
      tornado_effect_r2 <- 0
    }
    
    if(tornado_r8 == 1){
      tornado_effect_r8 <- sample(tornado_density_r8$x, 1, prob = tornado_density_r8$y)
    } else {
      tornado_effect_r8 <- 0
    }
    

    # Define starting inventory value
    start_inv_r9 <- df$inventory.r9[df$year == (y-1)]
    
    # Calculate the new growth, removal, mortality percent of inventory for the year
    df$new_growth_pct_r9[df$year == y] <- (y - 2014)*growth_roc_r9 + growth_int_r9 
    #+ df$new_growth_pct_r9[df$year == (y-1)]*growth_lag_roc_r9
    new_mort_pct_r9 <- (y - 2014)*mortality_roc_r9 + mortality_int_r9
    new_remv_pct_r9 <- (y - 2014)*removal_roc_r9 + removal_int_r9
    
    if(df$new_growth_pct_r9[df$year == y] < 0){
      df$new_growth_pct_r9[df$year == y] <- 0
    }
    if(new_mort_pct_r9 < 0){
      new_mort_pct_r9 <- 0
    }
    if(new_remv_pct_r9 < 0){
      new_remv_pct_r9 <- 0
    }
    
    # Calculate the inventory value for the year
    df$inventory.r9[df$year == y] <- start_inv_r9[[1]] + 
    df$new_growth_pct_r9[df$year == y] * start_inv_r9[[1]] - 
    new_mort_pct_r9 * start_inv_r9[[1]] - 
    new_remv_pct_r9 * start_inv_r9[[1]] - 
    fire_effect_r9 * start_inv_r9[[1]] -
    pest_effect_r9 * start_inv_r9[[1]] - 
      tornado_effect_r9 * start_inv_r9[[1]]
    
    # Define starting inventory value
    start_inv_r8 <- df$inventory.r8[df$year == (y-1)]
    
    # Calculate the new growth, removal, mortality percent of inventory for the year
    df$new_growth_pct_r8[df$year == y] <- (y - 2014)*growth_roc_r8 + growth_int_r8 
    #+ df$new_growth_pct_r8[df$year == (y-1)]*growth_lag_roc_r8
    df$new_mort_pct_r8[df$year == y] <- (y - 2014)*mortality_roc_r8 + mortality_int_r8 
    #+ df$new_mort_pct_r8[df$year == (y - 1)]*mortality_lag_roc_r8
    df$new_remv_pct_r8[df$year == y] <- (y - 2014)*removal_roc_r8 + removal_int_r8 
    #+ df$new_remv_pct_r8[df$year == (y-1)]*removal_lag_roc_r8

    if(df$new_growth_pct_r8 < 0){
      df$new_growth_pct_r8 <- 0
    }
    if(df$new_mort_pct_r8 < 0){
      df$new_mort_pct_r8 <- 0
    }
    if(df$new_remv_pct_r8 < 0){
      df$new_remv_pct_r8 <- 0
    }    
        
    # Calculate the inventory value for the year
    df$inventory.r8[df$year == y] <- start_inv_r8[[1]] + 
    df$new_growth_pct_r8[df$year == y] * start_inv_r8[[1]] - 
    df$new_mort_pct_r8[df$year == y] * start_inv_r8[[1]] - 
    df$new_remv_pct_r8[df$year == y] * start_inv_r8[[1]] - 
    fire_effect_r8 * start_inv_r8[[1]]-
    pest_effect_r8 * start_inv_r8[[1]] - 
      tornado_effect_r8 * start_inv_r8[[1]]
    
    # Define starting inventory value
    start_inv_r2 <- df$inventory.r2[df$year == (y-1)]
    
    # Calculate the new growth, removal, mortality percent of inventory for the year
    new_growth_pct_r2 <- (y - 2014)*growth_roc_r2 + growth_int_r2
    df$new_mort_pct_r2[df$year == y] <- (y - 2014)*mortality_roc_r2 + mortality_int_r2 
    #+ df$new_mort_pct_r2[df$year == (y-1)]*mortality_lag_roc_r2
    new_remv_pct_r2 <- (y - 2014)*removal_roc_r2 + removal_int_r2
    
    if(new_growth_pct_r2 < 0){
      new_growth_pct_r2 <- 0
    }
    if(df$new_mort_pct_r2[df$year == y] < 0){
      df$new_mort_pct_r2[df$year == y] <- 0
    }
    if(new_remv_pct_r2 < 0){
      new_remv_pct_r2 <- 0
    }    
    
    # Calculate the inventory value for the year
    df$inventory.r2[df$year == y] <- start_inv_r2[[1]] + 
    new_growth_pct_r2 * start_inv_r2[[1]] - 
    df$new_mort_pct_r2[df$year == y] * start_inv_r2[[1]] - 
    new_remv_pct_r2 * start_inv_r2[[1]] - 
    fire_effect_r2 * start_inv_r2[[1]] -  
    pest_effect_r2 * start_inv_r2[[1]] - 
      tornado_effect_r2 * start_inv_r2[[1]]
    
    y <- y+1
  }
  
  df$total_inventory = df$inventory.r9 + df$inventory.r8 + df$inventory.r2
  return(df$total_inventory)
}
```
```{r}
# This is where we define the number of times that we want the simulation to run
new_df <- tibble(years = years)
for (i in 2:10) {
  new_df[, i] <- run_sim()
}
```
```{r}
new_df_long <- new_df %>% gather(key = sim_number, value = value, - years)
```
```{r}
# Plot the results
ggplot(new_df_long, aes(x = years, y = value, colour = sim_number)) + 
  geom_smooth() +
  scale_y_continuous(limits = c(0, 85000000000)) +
  theme_minimal() +
  theme(legend.position = "none") +
  stat_summary(fun.y=mean, geom="line", colour="black", size = 2) +
  ggtitle("White Oak Forecast Simulation", subtitle = "The black line indicates the average forecasted value for each year.") +
  labs(y = "Inventory", x = "Year")
```


### Next Steps

Next we need to get a better understanding of the climate scenarios and how those may affect the growth/removal/mortality of the white oak population. Once we get an understanding of the effect, we can simulate the future population under each of these scenarios to get a more realistic prediction for white oak.

```{r echo = FALSE}

# simple_mod <- lm(total_inventory ~ growth.pct + remv.pct + mort.pct + total_acres, data = mod.dat.test)
# summary(simple_mod)
# 
# # multicollinearity
# vif(simple_mod)
# 
# # non constant variance
# ncvTest(simple_mod)
# 
# # independence of errors
# durbinWatsonTest(simple_mod)
# 
# 
# 
# simple_mod2 <- lm(total_inventory ~ total_growth + total_removal + total_mortality + total_acres, data = mod.dat.test)
# summary(simple_mod2)
# 
# vif(simple_mod2)
# ncvTest(simple_mod2)
# durbinWatsonTest(simple_mod2)

# Build linear models for each and save summary stats
growth.mod.r9 <- lm(growth.pct ~ EVAL_GRP_YEAR, data = mod.dat.r9)
growth.summary.r9 <- summary(growth.mod.r9)
mort.mod.r9 <- lm(mort.pct ~ EVAL_GRP_YEAR, data = mod.dat.r9)
mort.summary.r9 <- summary(mort.mod.r9)
remv.mod.r9 <- lm(log(remv.pct) ~ EVAL_GRP_YEAR, data = mod.dat.r9)
remv.summary.r9 <- summary(remv.mod.r9)

growth.mod.r8 <- lm(growth.pct ~ EVAL_GRP_YEAR, data = mod.dat.r8)
growth.summary.r8 <- summary(growth.mod.r8)
mort.mod.r8 <- lm(mort.pct ~ EVAL_GRP_YEAR, data = mod.dat.r8)
mort.summary.r8 <- summary(mort.mod.r8)
remv.mod.r8 <- lm(log(remv.pct) ~ EVAL_GRP_YEAR, data = mod.dat.r8)
remv.summary.r8 <- summary(remv.mod.r8)

growth.mod.r2 <- lm(growth.pct ~ EVAL_GRP_YEAR, data = mod.dat.r2)
growth.summary.r2 <- summary(growth.mod.r2)
mort.mod.r2 <- lm(mort.pct ~ EVAL_GRP_YEAR, data = mod.dat.r2)
mort.summary.r2 <- summary(mort.mod.r2)
remv.mod.r2 <- lm(remv.pct ~ EVAL_GRP_YEAR, data = mod.dat.r2)
remv.summary.r2 <- summary(remv.mod.r2)

library(lmtest)
par(mfrow = c(2, 2))
plot(remv.mod.r2)
shapiro.test(mod.dat.r2$remv.pct)
dwtest(remv.mod.r2)
library(gvlma)
summary(gvlma(remv.mod.r2))
# fail to reject the null
plot(remv.mod.r9)
dwtest(remv.mod.r9)
summary(gvlma(remv.mod.r9))
# fail to reject the null
plot(remv.mod.r8)
dwtest(remv.mod.r8)
summary(gvlma(remv.mod.r8))
##### REJECT THE NULL - how to deal with this?

dwtest(mort.mod.r2)
summary(gvlma(mort.mod.r2))
##### REJECT THE NULL - how to deal with this?
dwtest(mort.mod.r9)
summary(gvlma(mort.mod.r9))
# fail to reject the null
dwtest(mort.mod.r8)
summary(gvlma(mort.mod.r8))
##### REJECT THE NULL - how to deal with this?

dwtest(growth.mod.r2)
summary(gvlma(growth.mod.r2))
# fail to reject the null
dwtest(growth.mod.r9)
summary(gvlma(growth.mod.r9))
##### REJECT THE NULL - how to deal with this?
dwtest(growth.mod.r8)
summary(gvlma(mort.mod.r8))
##### REJECT THE NULL - how to deal with this?



mean(growth.mod.r9$residuals)
plot(fitted(growth.mod.r9), residuals(growth.mod.r9))


library(tseries)

ts_growth_r9 <- ts(mod.dat.r9$growth.pct, start = 2003, end = 2014)
ar_growth_r9 <- ar.ols(ts_growth_r9)


ggAcf(mod.dat.r9$growth.pct)
```

Documentation
=======================================================================

#### *Conclusion/Result*

**The white oak population will stay relatively flat in the near term.** Due to limited historical data, it is important to use caution when interpreting these results further into the forecast horizon. With only 13 annual data points to build the forecast, looking out past ten years serves best as a directional indication of the White Oak population.

#### *Methodology*

To build a forecast for the white oak population over the next 50 years, we decided to use a **simulation**. A simulation will allow us to forecast current trends in growth, removal, and mortality, as well as incorporate variables to capture the rare events such as catastrophic wildfires, category 5 tornados, and large-scale pest outbreaks. These are events that we would not expect to occur every year, or even once every 10 years. Therefore, by incorporating the probability of these events to occur each year we are able to get a better understanding of how potential environmental disasters can impact the population.

Included in the simulation is the current trend of growth, removal, and mortality, as well as the associated error so that we can be sure to capture the full range of where we might expect the population to be. These trends, as well as the potential for environmental outbreaks, will give us a final forecast that will be based off the average of all the simulated outcomes, and that should be relatively close to where we would expect the true population to be over the next few years.

#### *Purpose*

The Brown-Forman production team has requested a forecast of the White Oak population in order to better understand any short and/or long term risk associated with the population growth and mortality. A previous consultant had provided a forecast that had some fairly significant errors, and our goal was to enhance the consultant's work so that the resulting forecast is up-to-date and accurate.

#### *Data*

The data used for this analysis was received from the U.S. Forest Service. The data contains annual, unit level volumes for White Oak on private, non-industrial lands in the Eastern states, and includes detail on the Growing Stock Inventory, Growth, Removal, and Mortality in cubic feet. The data spans the period 2013 through 2014.

#### *Simulation Dashboard*

In order to forecast the White Oak population, we first calculated the yearly percent of inventory for the growth, removal, and mortality volumes. We then built a simple linear regression model to describe the trend of the growth, removal, and mortality, individually. Using the slope coefficient and an intercept based off the average of the percent of inventory for 2010 to 2014, we predicted the growth, removal, and mortality percent of inventory for the years 2015 through 2064. Using these values, as well as user-input values for annual urbanization percent of inventory, we calculated a forecast for the White Oak inventory.

As well as an inventory forecast, we also provide an upper bound and a lower bound (95% confidence interval), based off of the upper and lower estimates for the slope coefficient from the model. The 95% confidence interval means that we are 95% confident that the true values lies somewhere between the upper and lower bounds.