---
title: "White Oak Population Forecast"
resource_files:
- dashboard/states-shape/cb_2016_us_state_500k.shp
- dashboard/unit_bnd/eastunits.shp
- cb_2016_us_state_500k.shp
- eastunits.shp
#- cb_2016_us_state_500k.cpg
#- cb_2016_us_state_500k.dbf
#- cb_2016_us_state_500k.prj
#- cb_2016_us_state_500k.shp
#- cb_2016_us_state_500k.shp.ea.iso.xml
#- cb_2016_us_state_500k.shp.iso.xml
#- cb_2016_us_state_500k.shp.xml
#- cb_2016_us_state_500k.shx
#- eastunits.dbf
#- eastunits.shp
#- eastunits.shp.xml
#- eastunits.shx
- national_forestry_service_regions.cpg
- national_forestry_service_regions.dbf
- national_forestry_service_regions.prj
- national_forestry_service_regions.qpj 
- national_forestry_service_regions.shp 
- national_forestry_service_regions.shx
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    logo: brown-forman_416x416.jpg
    orientation: rows
    source_code: embed
    theme: readable
    vertical_layout: fill
---

```{r}
# Change the color of the title bar and the input sidebar. Remove the background from the leaflet map.

```

 <style>                     
.navbar {
  background-color: #C0C0C0;
}

.section.sidebar {
  background-color: #DCDCDC;
}

.leaflet-container {
    background-color:rgba(255,0,0,0.0);
}
</style> 

```{r setup, include=FALSE}
# Load required libraries
library(flexdashboard)
library(tidyverse)
library(ggplot2)
library(shiny)
library(dygraphs)
library(plotly)
library(formattable)
library(leaflet)
library(rgdal)

```

Simulation
=======================================================================
```{r}
# Read in state shape file
states_shp <- readOGR("MapFiles", layer = "national_forestry_service_regions",
                      GDAL1_integer64_policy = TRUE, verbose = FALSE)






```



Documentation
=======================================================================

#### Data

The data used for this analysis was received from the U.S. Forest Service. The data contains annual, unit level volumes for White Oak on private, non-industrial lands in the Eastern states, and includes detail on the Growing Stock Inventory, Growth, Removal, and Mortality in cubic feet. The data spans the period 1991 through 2014.

#### Method

##### *Forecast*

In order to forecast the White Oak population, we first calculated the yearly percent of inventory for the growth, removal, and mortality volumes. We then built a simple linear regression model to describe the trend of the growth, removal, and mortality, individually. Using the slope coefficient and an intercept based off the average of the percent of inventory for 2010 to 2014, we predicted the growth, removal, and mortality percent of inventory for the years 2015 through 2064. Using these values, as well as user-input values for annual urbanization percent of inventory, we calculated a forecast for the White Oak inventory.

As well as an inventory forecast, we also provide an upper bound and a lower bound (95% confidence interval), based off of the upper and lower estimates for the slope coefficient from the model. The 95% confidence interval means that we are 95% confident that the true values lies somewhere between the upper and lower bounds.

##### *Alternate Forecast*

We used the same methods to compile the alternate forecast, but instead of forecasting removal, we leave it as a user-defined field, similar to how the previous consultants compiled the forecast. 

However, we believe that predicting removal in the same manner as growth and mortality provides a better picture of the forecast, as the historic data could provide a better indication of the future trend rather than leaving it to the user to define.

#### Comments

Since we are only using 23 annual data points to build this forecast, it is important to use caution when interpreting these results further into the forecast horizon. As you look out past ten years, the forecasts serve best as a directional indication of the White Oak population.

The state forecast is only provided through 2040 due to a lack of data at the state level (the forecast for some states is based off of less than 10 years of data).