---
title: "Simulation Concept"
output:
  html_document:
    df_print: paged
---

```{r echo = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.width = 9)

library(tidyverse)
library(ggplot2)
library(readr)
library(tibbletime)
library(corrplot)
library(utf8)
library(stringr)
library(lmtest)
library(car)
library(gvlma)
library(sandwich)
library(forecast)
```

```{r echo = FALSE}
# Read in White Oak inventory files from USFS
inv23 <- read_csv("WhiteOakData/white_oak_inv_23.csv")
inv24 <- read_csv("WhiteOakData/white_oak_inv_24.csv")
inv33 <- read_csv("WhiteOakData/white_oak_inv_33.csv")

inventory <- rbind(inv23, inv24, inv33)
inventory[is.na(inventory)] <- 0

# Create unique identifier for UNITS
inventory$UNIT <- paste0(inventory$STATECD, "0", inventory$UNITCD)

# Read in the Growth, Removal, Mortality data cleaned and aggregated in the for-dashboard script
oak <- read_csv("WhiteOakData/white-oak-clean.csv")
oak[is.na(oak)] <- 0

# Summarise the necessary columns to prepare for merging with the inventory data
oak_merge <- oak %>%
  group_by(STATE_ABBR, EVAL_GRP_YEAR, UNIT) %>%
  summarise(total_removal = sum(SELECT_WHITE_OAK_REMOVAL, na.rm = TRUE),
            total_growth = sum(SELECT_WHITE_OAK_GROWTH, na.rm = TRUE),
            total_mortality = sum(SELECT_WHITE_OAK_MORTALITY, na.rm = TRUE)) %>%
  # total growth is actually net growth, so to get the actual growth we need to add back in mortality
  mutate(total_growth_true = total_growth + total_mortality)

oak_merge$total_growth <- NULL
colnames(oak_merge)[6] <- "total_growth"

# Summarise the Select WO and Total Hardwood Growing Stock (GS) Cubic Feet (CUFT) to prepare for mergin with GRM data
inventory_merge <- inventory %>%
  group_by(STATE_ABBR, EVAL_GRP_YEAR, UNIT) %>%
  summarise(total_inventory = sum(SELECT_WHITE_OAK_INV_CUFT_GS, na.rm = TRUE),
            total_hardwood = sum(TOTAL_HARDWOOD_INV_CUFT_GS, na.rm = TRUE))

oak_total <- merge(oak_merge, inventory_merge, by = c("STATE_ABBR", "EVAL_GRP_YEAR", "UNIT"))

# Create a df summarised to the state level
mod.dat <- oak_total %>%
  group_by(STATE_ABBR, EVAL_GRP_YEAR) %>%
  summarise(total_removal = sum(total_removal),
            total_growth = sum(total_growth),
            total_mortality = sum(total_mortality),
            total_inventory = sum(total_inventory),
            total_hardwood = sum(total_hardwood))

# Hardcode FS regions
r9 <- c("MN", "IA", "MO", "WI", "IL", "MI", "IN", "OH", "WV", "PA", "NJ", "DE", "MD", "NY", "CT", "RI", "MA", "VT", "NH", "ME")

r2 <- c("ND", "SD", "NE", "KS")

r8 <- c("TX", "OK", "LA", "AR", "MS", "AL", "GA", "FL", "TN", "KY", "VA", "NC", "SC")

# ifelse to prepare mod.dat for separation by region
mod.dat$region <- ifelse(mod.dat$STATE_ABBR %in% r9, "R9",
                         ifelse(mod.dat$STATE_ABBR %in% r8, "R8",
                                ifelse(mod.dat$STATE_ABBR %in% r2, "R2", "NA")))

# write_csv in case anything bad happens later
write_csv(mod.dat, "mod_dat_final.csv")

# Read in file for state abbreviations
pdsi_states <- read_csv("pdsi_states.csv")

# Read in invasive species/pest data
pest_output <- read_csv("pest_output.csv")

# merge with state df to get abbreviations
pest_merge <- merge(pest_output, pdsi_states, by = "State")
colnames(pest_merge)[5] <- "STATE_ABBR"

# add regions
pest_merge$region <- ifelse(pest_merge$STATE_ABBR %in% r9, "R9",
                         ifelse(pest_merge$STATE_ABBR %in% r8, "R8",
                                ifelse(pest_merge$STATE_ABBR %in% r2, "R2", "NA")))

# read in data for the total land area in each state
state_land_area <- read_csv("state_land_area.csv")

# Fix some of the columns
state_land_area <- separate(data = state_land_area, col = `State`, into = c("State", "Other"), sep = " flag")
state_land_area <- separate(data = state_land_area, col = `Land Area (sq mi)`, into = c("land_area", "other"), sep = "\\.")
state_land_area <- state_land_area %>%
  select(State, land_area)

state_land_area$land_area <- gsub(",", "", state_land_area$land_area)
state_land_area$land_area2 <- as.numeric(state_land_area$land_area)
state_land_area$land_area <- state_land_area$land_area2
state_land_area$land_area2 <- NULL

# merge with state abbreviation file
state_land_area_merge <- merge(state_land_area, pdsi_states, by = c("State")) %>%
  mutate(land_area_acres = land_area*640)

state_land_area_merge$land_area <- NULL
state_land_area_merge$State <- NULL

colnames(state_land_area_merge) <- c("STATE_ABBR", "land_area_acres")

# determine regions
state_land_area_merge$region <- ifelse(state_land_area_merge$STATE_ABBR %in% r9, "R9",
                         ifelse(state_land_area_merge$STATE_ABBR %in% r8, "R8",
                                ifelse(state_land_area_merge$STATE_ABBR %in% r2, "R2", "NA")))

# summarise to get total land area by region
state_land_area_reg <- state_land_area_merge %>%
  group_by(region) %>%
  summarise(land_area_acres = sum(land_area_acres))

# filter pest data by region

pest_r8 <- pest_merge %>% 
  filter(region == "R8") %>% 
  group_by(Year) %>% 
  summarise(total_acres = sum(Acres))

pest_r9 <- pest_merge %>% 
  filter(region == "R9") %>% 
  group_by(Year) %>% 
  summarise(total_acres = sum(Acres))

pest_r2 <- pest_merge %>% 
  filter(region == "R2") %>% 
  group_by(Year) %>% 
  summarise(total_acres = sum(Acres))


# How many times was the pest volume in each region greater than the region's average? (a potential way to quatify the outbreaks)

pest_r9$outbreak <- ifelse(pest_r9$total_acres > mean(pest_r9$total_acres), 1, 0)
pest_r8$outbreak <- ifelse(pest_r8$total_acres > mean(pest_r8$total_acres), 1, 0)
pest_r2$outbreak <- ifelse(pest_r2$total_acres > mean(pest_r2$total_acres), 1, 0)

# determine the % of land area destroyed by pests
pest_r2$reg_acres <- state_land_area_reg$land_area_acres[state_land_area_reg$region == "R2"]
pest_r9$reg_acres <- state_land_area_reg$land_area_acres[state_land_area_reg$region == "R9"]
pest_r8$reg_acres <- state_land_area_reg$land_area_acres[state_land_area_reg$region == "R8"]

pest_r2$pct_acres <- pest_r2$total_acres/pest_r2$reg_acres
pest_r9$pct_acres <- pest_r9$total_acres/pest_r9$reg_acres
pest_r8$pct_acres <- pest_r8$total_acres/pest_r8$reg_acres

# How often were there category 5 tornados? 
# read in tornado data
tornado_dat <- read_csv("Tornado/Consolidated_Tornados_50_16.csv")

colnames(tornado_dat)[2] <- "STATE_ABBR"

# determine regions
tornado_dat$region <- ifelse(tornado_dat$STATE_ABBR %in% r9, "R9",
                         ifelse(tornado_dat$STATE_ABBR %in% r8, "R8",
                                ifelse(tornado_dat$STATE_ABBR %in% r2, "R2", "NA")))

# filter to category 5
tornado_dat_5 <- tornado_dat %>% 
  filter(magnitude == 5) %>% 
  group_by(region, year) %>% 
  summarise(acres_destroyed = sum(acres))

# separate by region and determine % of land area destroyed
tornado_2 <- tornado_dat_5 %>% 
  filter(region == "R2") %>% 
  mutate(region_acres = state_land_area_reg$land_area_acres[state_land_area_reg$region == "R2"],
         pct_acres = acres_destroyed/region_acres)

tornado_9 <- tornado_dat_5 %>% 
  filter(region == "R9") %>% 
  mutate(region_acres = state_land_area_reg$land_area_acres[state_land_area_reg$region == "R9"],
         pct_acres = acres_destroyed/region_acres)

tornado_8 <- tornado_dat_5 %>% 
  filter(region == "R8") %>% 
  mutate(region_acres = state_land_area_reg$land_area_acres[state_land_area_reg$region == "R8"],
         pct_acres = acres_destroyed/region_acres)


# read in fire data
more_fire <- read_csv("more_fire_final.csv")

# since this data is from the FS, we will use total inventory to determine the % of land destroyed
fire_inv <- mod.dat  %>% 
  filter(EVAL_GRP_YEAR == 2014) %>% 
  group_by(EVAL_GRP_YEAR, region) %>% 
  summarise(total_inv = sum(total_inventory, na.rm = TRUE))

# separate by region
fire_r8_df <- more_fire %>% 
  filter(region == "R8") %>% 
  group_by(FIRE_YEAR) %>% 
  summarise(total_acres = sum(tot_acres))

fire_r9_df <- more_fire %>% 
  filter(region == "R9") %>% 
  group_by(FIRE_YEAR) %>% 
  summarise(total_acres = sum(tot_acres))

fire_r2_df <- more_fire %>% 
  filter(region == "R2") %>% 
  group_by(FIRE_YEAR) %>% 
  summarise(total_acres = sum(tot_acres))

# calculate % forest destroyed
fire_r2_df$reg_acres <- fire_inv$total_inv[fire_inv$region == "R2"]
fire_r9_df$reg_acres <- fire_inv$total_inv[fire_inv$region == "R9"]
fire_r8_df$reg_acres <- fire_inv$total_inv[fire_inv$region == "R8"]

fire_r2_df$pct_acres <- fire_r2_df$total_acres/fire_r2_df$reg_acres
fire_r9_df$pct_acres <- fire_r9_df$total_acres/fire_r9_df$reg_acres
fire_r8_df$pct_acres <- fire_r8_df$total_acres/fire_r8_df$reg_acres

# separate our GRM, inventory data by region for use in simulation
mod.dat.r9 <- mod.dat %>% 
  filter(region == "R9") %>% 
  group_by(EVAL_GRP_YEAR) %>% 
  summarise(total_inventory = sum(total_inventory),
            total_mortality = sum(total_mortality),
            total_removal = sum(as.numeric(total_removal)),
            total_growth = sum(as.numeric(total_growth)))

mod.dat.r8 <- mod.dat %>% 
  filter(region == "R8") %>% 
  group_by(EVAL_GRP_YEAR) %>% 
  summarise(total_inventory = sum(total_inventory),
            total_mortality = sum(total_mortality),
            total_removal = sum(as.numeric(total_removal)),
            total_growth = sum(as.numeric(total_growth)))

mod.dat.r2 <- mod.dat %>% 
  filter(region == "R2") %>% 
  group_by(EVAL_GRP_YEAR) %>% 
  summarise(total_inventory = sum(total_inventory),
            total_mortality = sum(total_mortality),
            total_removal = sum(as.numeric(total_removal)),
            total_growth = sum(as.numeric(total_growth)))


```

In the simulation outline below, we sample from a normal distribution with a mean equal to the coefficient of a linear model and a standard deviation equal to the standard error on the coefficient estimate in order to obtain rate of change values the growth, removal, and mortality percent of inventory. We then calculate the new percent of inventory each year by multiplying the number of years elapsed by the rate of chance (ROC) and add an intercept (calculated by taking the mean of the last 5 years of data). We also include a yearly shock value to provide a simple example of how we can incorporate the outside variables like wildfire, pests, drought, etc. The initial inventory value is the true inventory in the year 2014, and each subsequent year is calculated by adding/subtracting the growth, removal, mortality, and yearly shock values.

```{r}

# Calculate growth, removal, mortality percent of inventory
mod.dat.r9 <- mod.dat.r9 %>% 
  mutate(growth.pct = total_growth/total_inventory,
         remv.pct = total_removal/total_inventory, 
         mort.pct = total_mortality/total_inventory)

mod.dat.r8 <- mod.dat.r8 %>% 
  mutate(growth.pct = total_growth/total_inventory,
         remv.pct = total_removal/total_inventory, 
         mort.pct = total_mortality/total_inventory)

mod.dat.r2 <- mod.dat.r2 %>% 
  mutate(growth.pct = total_growth/total_inventory,
         remv.pct = total_removal/total_inventory, 
         mort.pct = total_mortality/total_inventory)

# Filter to 2003 to 2014 for consistency in measurement method

mod.dat.r9 <- filter(mod.dat.r9, EVAL_GRP_YEAR >= 2003 & EVAL_GRP_YEAR <= 2014)
mod.dat.r8 <- filter(mod.dat.r8, EVAL_GRP_YEAR >= 2003 & EVAL_GRP_YEAR <= 2014)
mod.dat.r2 <- filter(mod.dat.r2, EVAL_GRP_YEAR >= 2003 & EVAL_GRP_YEAR <= 2014)

# Build linear models for each and save summary stats
growth.mod.r9 <- lm(growth.pct ~ EVAL_GRP_YEAR, data = mod.dat.r9)
growth.summary.r9 <- summary(growth.mod.r9)
mort.mod.r9 <- lm(mort.pct ~ EVAL_GRP_YEAR, data = mod.dat.r9)
mort.summary.r9 <- summary(mort.mod.r9)
remv.mod.r9 <- lm(remv.pct ~ EVAL_GRP_YEAR, data = mod.dat.r9)
remv.summary.r9 <- summary(remv.mod.r9)

growth.mod.r8 <- lm(growth.pct ~ EVAL_GRP_YEAR, data = mod.dat.r8)
growth.summary.r8 <- summary(growth.mod.r8)
mort.mod.r8 <- lm(mort.pct ~ EVAL_GRP_YEAR, data = mod.dat.r8)
mort.summary.r8 <- summary(mort.mod.r8)
remv.mod.r8 <- lm(remv.pct ~ EVAL_GRP_YEAR, data = mod.dat.r8)
remv.summary.r8 <- summary(remv.mod.r8)

growth.mod.r2 <- lm(growth.pct ~ EVAL_GRP_YEAR, data = mod.dat.r2)
growth.summary.r2 <- summary(growth.mod.r2)
mort.mod.r2 <- lm(mort.pct ~ EVAL_GRP_YEAR, data = mod.dat.r2)
mort.summary.r2 <- summary(mort.mod.r2)
remv.mod.r2 <- lm(remv.pct ~ EVAL_GRP_YEAR, data = mod.dat.r2)
remv.summary.r2 <- summary(remv.mod.r2)

# Check Assumptions

# Independence of Errors (Autocorrelation)

# Breusch Godfrey test
bgtest(remv.mod.r2) # fail to reject
bgtest(remv.mod.r9) # fail to reject
bgtest(remv.mod.r8) # fail to reject
bgtest(mort.mod.r2) # fail to reject
bgtest(mort.mod.r9) # fail to reject
bgtest(mort.mod.r8) # fail to reject
bgtest(growth.mod.r2) # fail to reject
bgtest(growth.mod.r9) # fail to reject
bgtest(growth.mod.r8) # fail to reject

# Ljung Box Test
Box.test(remv.mod.r2$residuals, type="Ljung-Box") # fail to reject
Box.test(remv.mod.r9$residuals, type="Ljung-Box") # fail to reject
Box.test(remv.mod.r8$residuals, type="Ljung-Box") # fail to reject
Box.test(mort.mod.r2$residuals, type="Ljung-Box") # fail to reject
Box.test(mort.mod.r9$residuals, type="Ljung-Box") # fail to reject
Box.test(mort.mod.r8$residuals, type="Ljung-Box") # fail to reject
Box.test(growth.mod.r2$residuals, type="Ljung-Box") # fail to reject
Box.test(growth.mod.r9$residuals, type="Ljung-Box") # fail to reject
Box.test(growth.mod.r8$residuals, type="Ljung-Box") # fail to reject

# Heteroscedasticity

# Breusch Pagan test

bptest(growth.mod.r9) # fail to reject
bptest(growth.mod.r8) # reject the null
bptest(growth.mod.r2) # fail to reject
bptest(mort.mod.r9) # fail to reject
bptest(mort.mod.r8) # reject the null
bptest(mort.mod.r2) # fail to reject 
bptest(remv.mod.r9) # fail to reject
bptest(remv.mod.r8) # reject the null
bptest(remv.mod.r2) # reject the null

# Normality
shapiro.test(growth.mod.r9$residuals) # fail to reject
shapiro.test(growth.mod.r8$residuals) # fail to reject
shapiro.test(growth.mod.r2$residuals) # fail to reject
shapiro.test(mort.mod.r9$residuals) # fail to reject
shapiro.test(mort.mod.r8$residuals) # fail to reject
shapiro.test(mort.mod.r2$residuals) # fail to reject 
shapiro.test(remv.mod.r9$residuals) # fail to reject
shapiro.test(remv.mod.r8$residuals) # fail to reject
shapiro.test(remv.mod.r2$residuals) # fail to reject

# Mean of errors is 0
mean(growth.mod.r9$residuals) # true
mean(growth.mod.r8$residuals) # true
mean(growth.mod.r2$residuals) # true
mean(mort.mod.r9$residuals) # true
mean(mort.mod.r8$residuals) # true
mean(mort.mod.r2$residuals) # true
mean(remv.mod.r9$residuals) # true
mean(remv.mod.r8$residuals) # true
mean(remv.mod.r2$residuals) # true

# Check Residuals




# Fix heteroscedasticity problems
remv.mod.r8$newse <- vcovHC(remv.mod.r8, type="HC3")
remv.summary.r8 <- coeftest(remv.mod.r8, remv.mod.r8$newse)
remv.mod.r2$newse <- vcovHC(remv.mod.r2, type="HC3")
remv.summary.r2 <- coeftest(remv.mod.r2, remv.mod.r2$newse)
mort.mod.r8$newse <- vcovHC(mort.mod.r8, type="HC3")
mort.summary.r8 <- coeftest(mort.mod.r8, mort.mod.r8$newse)
growth.mod.r8$newse <- vcovHC(growth.mod.r8, type="HC3")
growth.summary.r8 <- coeftest(growth.mod.r8, growth.mod.r8$newse)

```

```{r}

years <- 2014:2064

# Build simulation function
run_sim <- function() {
  # Define forecast period
  inv <- 0
  df <- tibble(year = years, inventory.r9 = inv, inventory.r8 = inv, inventory.r2 = inv, total_inventory = inv, new_growth_pct_r9 = inv, new_growth_pct_r8 = inv, new_remv_pct_r8 = inv, new_mort_pct_r2 = inv, new_mort_pct_r8 = inv)
  
  # Initial inventory equal to 2014 inventory
  df$inventory.r9[df$year == 2014] <- mod.dat.r9$total_inventory[nrow(mod.dat.r9)]
  df$inventory.r8[df$year == 2014] <- mod.dat.r8$total_inventory[nrow(mod.dat.r8)]
  df$inventory.r2[df$year == 2014] <- mod.dat.r2$total_inventory[nrow(mod.dat.r2)]
  
  df$new_growth_pct_r8[df$year == 2014] <- mod.dat.r8$growth.pct[nrow(mod.dat.r8)]
  df$new_growth_pct_r9[df$year == 2014] <- mod.dat.r9$growth.pct[nrow(mod.dat.r9)]
  df$new_mort_pct_r8[df$year == 2014] <- mod.dat.r8$mort.pct[nrow(mod.dat.r8)]
  df$new_mort_pct_r2[df$year == 2014] <- mod.dat.r2$mort.pct[nrow(mod.dat.r2)]
  df$new_remv_pct_r8[df$year == 2014] <- mod.dat.r8$remv.pct[nrow(mod.dat.r8)]
  
  # Build the normal distributions for growth, mortality, removal
  growth_dist_r9 <- rnorm(100, growth.mod.r9$coefficients[[2]], growth.summary.r9$coefficients[[4]])
  mortality_dist_r9 <- rnorm(100, mort.mod.r9$coefficients[[2]], mort.summary.r9$coefficients[[4]])
  removal_dist_r9 <- rnorm(100, remv.mod.r9$coefficients[[2]], remv.summary.r9$coefficients[[4]])
  
  growth_dist_r8 <- rnorm(100, growth.mod.r8$coefficients[[2]], growth.summary.r8[[4]])
  mortality_dist_r8 <- rnorm(100, mort.mod.r8$coefficients[[2]], mort.summary.r8[[4]])
  removal_dist_r8 <- rnorm(100, remv.mod.r8$coefficients[[2]], remv.summary.r8[[4]])

  growth_dist_r2 <- rnorm(100, growth.mod.r2$coefficients[[2]], growth.summary.r2$coefficients[[4]])
  mortality_dist_r2 <- rnorm(100, mort.mod.r2$coefficients[[2]], mort.summary.r2$coefficients[[4]])
  removal_dist_r2 <- rnorm(100, remv.mod.r2$coefficients[[2]], remv.summary.r2[[4]])
  
  # Define rate of change by sampling from distributions created above. We want the rate of change to be the same over the entire forecasted time period, but different for each simulation. This is why we define it prior to the while loop below.
  growth_roc_r9 <- base::sample(growth_dist_r9, 1)
  mortality_roc_r9 <- base::sample(mortality_dist_r9, 1)
  removal_roc_r9 <- base::sample(removal_dist_r9, 1)
  
  growth_roc_r8 <- base::sample(growth_dist_r8, 1)
  mortality_roc_r8 <- base::sample(mortality_dist_r8, 1)
  removal_roc_r8 <- base::sample(removal_dist_r8, 1)
  
  growth_roc_r2 <- base::sample(growth_dist_r2, 1)
  mortality_roc_r2 <- base::sample(mortality_dist_r2, 1)
  removal_roc_r2 <- base::sample(removal_dist_r2, 1)
  
  y <- 2015
  
  # Define intercepts by taking the average of the last 5 years for growth, removal, mortality percent of inventory
  growth_int_r9 <- mean(mod.dat.r9$growth.pct[(nrow(mod.dat.r9)-5):nrow(mod.dat.r9)])
  mortality_int_r9 <- mean(mod.dat.r9$mort.pct[(nrow(mod.dat.r9)-5):nrow(mod.dat.r9)])
  removal_int_r9 <- mean(mod.dat.r9$remv.pct[(nrow(mod.dat.r9)-5):nrow(mod.dat.r9)])
  
  growth_int_r8 <- mean(mod.dat.r8$growth.pct[(nrow(mod.dat.r8)-5):nrow(mod.dat.r8)])
  mortality_int_r8 <- mean(mod.dat.r8$mort.pct[(nrow(mod.dat.r8)-5):nrow(mod.dat.r8)])
  removal_int_r8 <- mean(mod.dat.r8$remv.pct[(nrow(mod.dat.r8)-5):nrow(mod.dat.r8)])
  
  growth_int_r2 <- mean(mod.dat.r2$growth.pct[(nrow(mod.dat.r2)-5):nrow(mod.dat.r2)])
  mortality_int_r2 <- mean(mod.dat.r2$mort.pct[(nrow(mod.dat.r2)-5):nrow(mod.dat.r2)])
  removal_int_r2 <- mean(mod.dat.r2$remv.pct[(nrow(mod.dat.r2)-5):nrow(mod.dat.r2)])
  
  # Determine probability of catastrophic fire in each region
  
  fire_r8_prob <- nrow(fire_r8_df)/(2014-min(fire_r8_df$FIRE_YEAR))
  fire_r9_prob <- nrow(fire_r9_df)/(2014-min(fire_r9_df$FIRE_YEAR))
  fire_r2_prob <- nrow(fire_r2_df)/(2014-min(fire_r2_df$FIRE_YEAR))
  
  fire_density_r9 <- density(fire_r9_df$pct_acres)
  fire_density_r9 <- data.frame(fire_density_r9$x, fire_density_r9$y)
  colnames(fire_density_r9) <- c("x","y")
  fire_density_r9 <- filter(fire_density_r9, x >= 0)
  
  fire_density_r8 <- density(fire_r8_df$pct_acres)
  fire_density_r8 <- data.frame(fire_density_r8$x, fire_density_r8$y)
  colnames(fire_density_r8) <- c("x","y")
  fire_density_r8 <- filter(fire_density_r8, x >= 0)
  
  fire_density_r2 <- density(fire_r2_df$pct_acres)
  fire_density_r2 <- data.frame(fire_density_r2$x, fire_density_r2$y)
  colnames(fire_density_r2) <- c("x","y")
  fire_density_r2 <- filter(fire_density_r2, x >= 0)
  
  # Determine probability of larger than normal pest outbreak
  
  pest_r8_prob <- sum(pest_r8$outbreak)/(max(pest_r8$Year)-min(pest_r8$Year))
  pest_r9_prob <- sum(pest_r9$outbreak)/(max(pest_r9$Year)-min(pest_r9$Year))
  pest_r2_prob <- sum(pest_r2$outbreak)/(max(pest_r2$Year)-min(pest_r2$Year))
  
  pest_density_r2 <- density(pest_r2$pct_acres[pest_r2$outbreak==1])
  pest_density_r2 <- data.frame(pest_density_r2$x, pest_density_r2$y)
  colnames(pest_density_r2) <- c("x","y")
  pest_density_r2 <- filter(pest_density_r2, x >= 0)
  
  pest_density_r8 <- density(pest_r8$pct_acres[pest_r8$outbreak==1])
  pest_density_r8 <- data.frame(pest_density_r8$x, pest_density_r8$y)
  colnames(pest_density_r8) <- c("x","y")
  pest_density_r8 <- filter(pest_density_r8, x >= 0)
  
  pest_density_r9 <- density(pest_r9$pct_acres[pest_r9$outbreak==1])
  pest_density_r9 <- data.frame(pest_density_r9$x, pest_density_r9$y)
  colnames(pest_density_r9) <- c("x","y")
  pest_density_r9 <- filter(pest_density_r9, x >= 0)
  
    # Determine probability of category 5 tornado
  
  tornado_r8_prob <- nrow(tornado_8)/(max(tornado_8$year)-min(tornado_8$year))
  tornado_r9_prob <- nrow(tornado_9)/(max(tornado_9$year)-min(tornado_9$year))
  tornado_r2_prob <- nrow(tornado_2)/(max(tornado_2$year)-min(tornado_2$year))
  
  tornado_density_r2 <- density(tornado_2$pct_acres)
  tornado_density_r2 <- data.frame(tornado_density_r2$x, tornado_density_r2$y)
  colnames(tornado_density_r2) <- c("x","y")
  tornado_density_r2 <- filter(tornado_density_r2, x >= 0)
  
  tornado_density_r8 <- density(tornado_8$pct_acres)
  tornado_density_r8 <- data.frame(tornado_density_r8$x, tornado_density_r8$y)
  colnames(tornado_density_r8) <- c("x","y")
  tornado_density_r8 <- filter(tornado_density_r8, x >= 0)
  
  tornado_density_r9 <- density(tornado_9$pct_acres)
  tornado_density_r9 <- data.frame(tornado_density_r9$x, tornado_density_r9$y)
  colnames(tornado_density_r9) <- c("x","y")
  tornado_density_r9 <- filter(tornado_density_r9, x >= 0)
  
  
  # Use while loop to build forecast year by year
  while (y < 2065) {
    
    # Will there be a fire, and if yes, what will the effect be?
    
    fire_r9 <- sample(c(0,1),1,prob = c((1-fire_r9_prob),fire_r9_prob))
    fire_r8 <- sample(c(0,1),1,prob = c((1-fire_r8_prob),fire_r8_prob))
    fire_r2 <- sample(c(0,1),1,prob = c((1-fire_r2_prob),fire_r2_prob))
    
    if(fire_r9 == 1){
      fire_effect_r9 <- sample(fire_density_r9$x, 1, prob = fire_density_r9$y)
    } else {
      fire_effect_r9 <- 0
    }
    
    if(fire_r8 == 1){
      fire_effect_r8 <- sample(fire_density_r8$x, 1, prob = fire_density_r8$y)
    } else {
      fire_effect_r8 <- 0
    }
    
    if(fire_r2 == 1){
      fire_effect_r2 <- sample(fire_density_r2$x, 1, prob = fire_density_r2$y)
    } else {
      fire_effect_r2 <- 0
    }
    
    # Will there be a pest outbreak? And if yes, what will the effect be?
    
    pest_ob_r9 <- sample(c(0,1),1,prob = c((1-pest_r9_prob),pest_r9_prob))
    pest_ob_r8 <- sample(c(0,1),1,prob = c((1-pest_r8_prob),pest_r8_prob))
    pest_ob_r2 <- sample(c(0,1),1,prob = c((1-pest_r2_prob),pest_r2_prob))
    
    if(pest_ob_r9 == 1){
      pest_effect_r9 <- sample(pest_density_r9$x, 1, prob = pest_density_r9$y)
    } else {
      pest_effect_r9 <- 0
    }
    
    if(pest_ob_r2 == 1){
      pest_effect_r2 <- sample(pest_density_r2$x, 1, prob = pest_density_r2$y)
    } else {
      pest_effect_r2 <- 0
    }
    
    if(pest_ob_r8 == 1){
      pest_effect_r8 <- sample(pest_density_r8$x, 1, prob = pest_density_r8$y)
    } else {
      pest_effect_r8 <- 0
    }
    
    # Will there be a big tornado outbreak? And if yes, what will the effect be?
    
    tornado_r9 <- sample(c(0,1),1,prob = c((1-tornado_r9_prob),tornado_r9_prob))
    tornado_r8 <- sample(c(0,1),1,prob = c((1-tornado_r8_prob),tornado_r8_prob))
    tornado_r2 <- sample(c(0,1),1,prob = c((1-tornado_r2_prob),tornado_r2_prob))
    
    if(tornado_r9 == 1){
      tornado_effect_r9 <- sample(tornado_density_r9$x, 1, prob = tornado_density_r9$y)
    } else {
      tornado_effect_r9 <- 0
    }
    
    if(tornado_r2 == 1){
      tornado_effect_r2 <- sample(tornado_density_r2$x, 1, prob = tornado_density_r2$y)
    } else {
      tornado_effect_r2 <- 0
    }
    
    if(tornado_r8 == 1){
      tornado_effect_r8 <- sample(tornado_density_r8$x, 1, prob = tornado_density_r8$y)
    } else {
      tornado_effect_r8 <- 0
    }
    

    # Define starting inventory value
    start_inv_r9 <- df$inventory.r9[df$year == (y-1)]
    
    # Calculate the new growth, removal, mortality percent of inventory for the year
    df$new_growth_pct_r9[df$year == y] <- (y - 2014)*growth_roc_r9 + growth_int_r9
    new_mort_pct_r9 <- (y - 2014)*mortality_roc_r9 + mortality_int_r9
    new_remv_pct_r9 <- (y - 2014)*removal_roc_r9 + removal_int_r9
    
    if(df$new_growth_pct_r9[df$year == y] < 0){
      df$new_growth_pct_r9[df$year == y] <- 0
    }
    if(new_mort_pct_r9 < 0){
      new_mort_pct_r9 <- 0
    }
    if(new_remv_pct_r9 < 0){
      new_remv_pct_r9 <- 0
    }
    
    # Calculate the inventory value for the year
    df$inventory.r9[df$year == y] <- start_inv_r9[[1]] + 
    df$new_growth_pct_r9[df$year == y] * start_inv_r9[[1]] - 
    new_mort_pct_r9 * start_inv_r9[[1]] - 
    new_remv_pct_r9 * start_inv_r9[[1]] - 
    fire_effect_r9 * start_inv_r9[[1]] -
    pest_effect_r9 * start_inv_r9[[1]] - 
      tornado_effect_r9 * start_inv_r9[[1]]
    
    # Define starting inventory value
    start_inv_r8 <- df$inventory.r8[df$year == (y-1)]
    
    # Calculate the new growth, removal, mortality percent of inventory for the year
    df$new_growth_pct_r8[df$year == y] <- (y - 2014)*growth_roc_r8 + growth_int_r8 
    df$new_mort_pct_r8[df$year == y] <- (y - 2014)*mortality_roc_r8 + mortality_int_r8 
    df$new_remv_pct_r8[df$year == y] <- (y - 2014)*removal_roc_r8 + removal_int_r8 

    if(df$new_growth_pct_r8 < 0){
      df$new_growth_pct_r8 <- 0
    }
    if(df$new_mort_pct_r8 < 0){
      df$new_mort_pct_r8 <- 0
    }
    if(df$new_remv_pct_r8 < 0){
      df$new_remv_pct_r8 <- 0
    }    
        
    # Calculate the inventory value for the year
    df$inventory.r8[df$year == y] <- start_inv_r8[[1]] + 
    df$new_growth_pct_r8[df$year == y] * start_inv_r8[[1]] - 
    df$new_mort_pct_r8[df$year == y] * start_inv_r8[[1]] - 
    df$new_remv_pct_r8[df$year == y] * start_inv_r8[[1]] - 
    fire_effect_r8 * start_inv_r8[[1]]-
    pest_effect_r8 * start_inv_r8[[1]] - 
      tornado_effect_r8 * start_inv_r8[[1]]
    
    # Define starting inventory value
    start_inv_r2 <- df$inventory.r2[df$year == (y-1)]
    
    # Calculate the new growth, removal, mortality percent of inventory for the year
    new_growth_pct_r2 <- (y - 2014)*growth_roc_r2 + growth_int_r2
    df$new_mort_pct_r2[df$year == y] <- (y - 2014)*mortality_roc_r2 + mortality_int_r2
    new_remv_pct_r2 <- (y - 2014)*removal_roc_r2 + removal_int_r2
    
    if(new_growth_pct_r2 < 0){
      new_growth_pct_r2 <- 0
    }
    if(df$new_mort_pct_r2[df$year == y] < 0){
      df$new_mort_pct_r2[df$year == y] <- 0
    }
    if(new_remv_pct_r2 < 0){
      new_remv_pct_r2 <- 0
    }    
    
    # Calculate the inventory value for the year
    df$inventory.r2[df$year == y] <- start_inv_r2[[1]] + 
    new_growth_pct_r2 * start_inv_r2[[1]] - 
    df$new_mort_pct_r2[df$year == y] * start_inv_r2[[1]] - 
    new_remv_pct_r2 * start_inv_r2[[1]] - 
    fire_effect_r2 * start_inv_r2[[1]] -  
    pest_effect_r2 * start_inv_r2[[1]] - 
      tornado_effect_r2 * start_inv_r2[[1]]
    
    y <- y+1
  }
  
  df$total_inventory = df$inventory.r9 + df$inventory.r8 + df$inventory.r2
  return(df$inventory.r9)
}
```
```{r warning = FALSE, message = FALSE}
# This is where we define the number of times that we want the simulation to run
new_df <- tibble(years = years)
for (i in 2:1501) {
  new_df[, i] <- run_sim()
}
```
```{r}
new_df_long <- new_df %>% gather(key = sim_number, value = value, - years)
med_df <- new_df_long %>% 
  group_by(years) %>% 
  summarise(value = mean(value)) %>% 
  mutate(sim_number = "mean") %>% 
  select(years, sim_number, value)
q90_df <- new_df_long %>% 
  group_by(years) %>% 
  summarise(value = quantile(value, 0.9)) %>% 
  mutate(sim_number = "q90") %>% 
  select(years, sim_number, value)
q10_df <- new_df_long %>% 
  group_by(years) %>% 
  summarise(value = quantile(value, 0.1)) %>% 
  mutate(sim_number = "q10") %>% 
  select(years, sim_number, value)
q975_df <- new_df_long %>% 
  group_by(years) %>% 
  summarise(value = quantile(value, 0.975)) %>% 
  mutate(sim_number = "q975") %>% 
  select(years, sim_number, value)
q025_df <- new_df_long %>% 
  group_by(years) %>% 
  summarise(value = quantile(value, 0.025)) %>% 
  mutate(sim_number = "q025") %>% 
  select(years, sim_number, value)
new_df_long <- rbind(med_df, q90_df, q10_df, q975_df, q025_df, new_df_long)

new_df_long$type <- ifelse(new_df_long$sim_number == "mean", "m", ifelse(grepl("q",new_df_long$sim_number)==TRUE, "q", "no"))
new_df_long$type <- factor(new_df_long$type)
#write_csv(new_df_long, "sim-for-dash-REGION-9.csv")
```


```{r}
# Plot the results
ggplot(new_df_long, aes(x = years, y = (value/1000000000), colour = type, group = sim_number, size = type)) + 
  geom_line() +
  scale_color_manual(values=c("lightgrey", "darkgreen", "black"))+
  scale_size_manual(values = c(.5, 2, 1))+
  scale_y_continuous(limits = c(0,100))+
  theme_minimal()+
  ggtitle("White Oak Forecast - Simulation", subtitle = "The black line indicates the mean forecast for each year.") +
  labs(y = "Inventory (Billions, Cubic Feet)", x = "Year")+ 
  theme(legend.position="none")
```


### Next Steps

Next we need to get a better understanding of the climate scenarios and how those may affect the growth/removal/mortality of the white oak population. Once we get an understanding of the effect, we can simulate the future population under each of these scenarios to get a more realistic prediction for white oak.

```{r echo = FALSE}

# simple_mod <- lm(total_inventory ~ growth.pct + remv.pct + mort.pct + total_acres, data = mod.dat.test)
# summary(simple_mod)
# 
# # multicollinearity
# vif(simple_mod)
# 
# # non constant variance
# ncvTest(simple_mod)
# 
# # independence of errors
# durbinWatsonTest(simple_mod)
# 
# 
# 
# simple_mod2 <- lm(total_inventory ~ total_growth + total_removal + total_mortality + total_acres, data = mod.dat.test)
# summary(simple_mod2)
# 
# vif(simple_mod2)
# ncvTest(simple_mod2)
# durbinWatsonTest(simple_mod2)

# Build linear models for each and save summary stats
growth.mod.r9 <- lm(growth.pct ~ EVAL_GRP_YEAR, data = mod.dat.r9)
growth.summary.r9 <- summary(growth.mod.r9)
mort.mod.r9 <- lm(mort.pct ~ EVAL_GRP_YEAR, data = mod.dat.r9)
mort.summary.r9 <- summary(mort.mod.r9)
remv.mod.r9 <- lm(log(remv.pct) ~ EVAL_GRP_YEAR, data = mod.dat.r9)
remv.summary.r9 <- summary(remv.mod.r9)

growth.mod.r8 <- lm(growth.pct ~ EVAL_GRP_YEAR, data = mod.dat.r8)
growth.summary.r8 <- summary(growth.mod.r8)
mort.mod.r8 <- lm(mort.pct ~ EVAL_GRP_YEAR, data = mod.dat.r8)
mort.summary.r8 <- summary(mort.mod.r8)
remv.mod.r8 <- lm(log(remv.pct) ~ EVAL_GRP_YEAR, data = mod.dat.r8)
remv.summary.r8 <- summary(remv.mod.r8)

growth.mod.r2 <- lm(growth.pct ~ EVAL_GRP_YEAR, data = mod.dat.r2)
growth.summary.r2 <- summary(growth.mod.r2)
mort.mod.r2 <- lm(mort.pct ~ EVAL_GRP_YEAR, data = mod.dat.r2)
mort.summary.r2 <- summary(mort.mod.r2)
remv.mod.r2 <- lm(remv.pct ~ EVAL_GRP_YEAR, data = mod.dat.r2)
remv.summary.r2 <- summary(remv.mod.r2)

library(lmtest)
par(mfrow = c(2, 2))
plot(remv.mod.r2)
shapiro.test(mod.dat.r2$remv.pct)
dwtest(remv.mod.r2)
library(gvlma)
summary(gvlma(remv.mod.r2))
# fail to reject the null
plot(remv.mod.r9)
dwtest(remv.mod.r9)
summary(gvlma(remv.mod.r9))
# fail to reject the null
plot(remv.mod.r8)
dwtest(remv.mod.r8)
summary(gvlma(remv.mod.r8))
##### REJECT THE NULL - how to deal with this?

dwtest(mort.mod.r2)
summary(gvlma(mort.mod.r2))
##### REJECT THE NULL - how to deal with this?
dwtest(mort.mod.r9)
summary(gvlma(mort.mod.r9))
# fail to reject the null
dwtest(mort.mod.r8)
summary(gvlma(mort.mod.r8))
##### REJECT THE NULL - how to deal with this?

dwtest(growth.mod.r2)
summary(gvlma(growth.mod.r2))
# fail to reject the null
dwtest(growth.mod.r9)
summary(gvlma(growth.mod.r9))
##### REJECT THE NULL - how to deal with this?
dwtest(growth.mod.r8)
summary(gvlma(mort.mod.r8))
##### REJECT THE NULL - how to deal with this?



mean(growth.mod.r9$residuals)
plot(fitted(growth.mod.r9), residuals(growth.mod.r9))


library(tseries)

ts_growth_r9 <- ts(mod.dat.r9$growth.pct, start = 2003, end = 2014)
ar_growth_r9 <- ar.ols(ts_growth_r9)


ggAcf(mod.dat.r9$growth.pct)
```