---
title: "White Oak Exploration"
author: "Paul Forst, Tammy Hang, Rachel Kopecky, Jack Letcher, Ian O'Connor"
output:
  html_document:
    theme: sandstone
    toc: yes
    toc_depth: 4
    toc_float: yes
  html_notebook:
    fig_caption: yes
    highlight: textmate
    theme: sandstone
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r echo = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 9)

library(tidyverse)
library(ggplot2)
library(readr)
library(tibbletime)
```

# Data

### White Oak

The White Oak data used for this analysis was received from the U.S. Forest Service (USFS). The data contains annual, unit level volumes for White Oak on private, non-industrial lands in the Eastern United States, and includes detail on the Growing Stock Inventory, Growth, Removal, and Mortality in cubic feet. The data spans the period 1968 through 2017.

### Climate

### Drought

### Urbanization

### Invasive Species

# Exploration

### White Oak

#### Inventory

The plot below shows the total White Oak inventory from 1968 through 2017, as recorded in the USFS database.

```{r}
# Read in and manipulate inventory data. These are the raw files from the US Forest service.
inv23 <- read_csv("white_oak_inv_23.csv")
inv24 <- read_csv("white_oak_inv_24.csv")
inv33 <- read_csv("white_oak_inv_33.csv")

inventory <- rbind(inv23, inv24, inv33)

inventory[is.na(inventory)] <- 0

# Create an identifier
inventory$UNIT <- paste0(inventory$STATECD, "0", inventory$UNITCD)

inventory %>% 
  group_by(EVAL_GRP_YEAR) %>% 
  summarise(total_inventory = sum(SELECT_WHITE_OAK_INV_CUFT_GS, na.rm = TRUE)) %>% 
  ggplot(aes(x = EVAL_GRP_YEAR, y = total_inventory)) +
  geom_line() +
  labs(title = "White Oak Inventory (1968 - 2017)",
       y = "White Oak, Cubic Feet",
       x = "Year") +
  theme_minimal()
```